# Regression I: K-nearest neighbours {#regression1}

## Overview 

This chapter continues our foray into answering predictive questions.
Here we will focus on *regression*, which is the process of predicting
*numerical* response variables.
This is unlike the past two chapters, which focused on predicting categorical
variables via classification. However, regression does have many similarities
to classification: for example, just as in the case of classification,
we will split our data into training, validation, and test sets, we will 
use `tidymodels` workflows, we will use a K-nearest neighbours 
approach to make predictions, and we will use cross-validation to choose K.
Because of how similar these procedures are, make sure to read Chapters
6 and 7 before reading this one---we will move a little bit faster here with the
concepts that have already been covered.
This chapter will primarily focus on the case where there is a single predictor and single
response variable of interest, but the end of the chapter shows how to perform
*multivariate regression*, i.e., regression with more than one predictor variable.

## Chapter learning objectives 
By the end of the chapter, students will be able to:

* Recognize situations where a simple regression analysis would be appropriate for making predictions.
* Explain the K-nearest neighbour (K-NN) regression algorithm and describe how it differs from K-NN classification.
* Interpret the output of a K-NN regression.
* In a dataset with two or more variables, perform K-nearest neighbour regression in R using a `tidymodels` workflow
* Execute cross-validation in R to choose the number of neighbours.
* Evaluate K-NN regression prediction accuracy in R using a test data set and an appropriate metric (*e.g.*, root means square prediction error).
* In the context of K-NN regression, compare and contrast goodness of fit and prediction properties (namely RMSE vs RMSPE).
* Describe the advantages and disadvantages of the K-nearest neighbour regression approach.

## The regression problem

Regression, like classification, is a predictive problem setting where we want
to use past information to predict future observations. But in the case of
regression, the goal is to predict *numerical* values instead of *categorical* values. 
For example, we could try to use the number of hours a person spends on
exercise each week to predict their race time in the annual Boston marathon. As 
another example, we could try to use the size of a house to
predict its sale price. Both of these response variables---race time and sale price---are 
numerical, and so predicting them given past data is considered a regression problem.

Just like in the classification setting, there are many possible methods that we can use 
to predict numerical response variables. In this chapter we will
focus on the **K-nearest neighbours** algorithm, and in the next chapter
we will study **linear regression**.
In your future studies, you might encounter regression trees, splines,
and general local regression methods; see the additional resources
section at the end of the next chapter for where to begin learning more about
these other methods. 

Many of the concepts from classification map over to the setting of regression. For example, 
a regression model predicts a new observation's response variable based on the response variables
for similar observations in the data set of past observations. When building a regression model,
we first split the data into training and test sets, in order to ensure that we assess the performance
of our method on observations not seen during training. And finally, we can use cross-validation to evaluate different
choices of model parameters (e.g., K in a K-nearest neighbours model). The major difference
is that we are now predicting numerical variables instead of categorical variables.

> You can usually tell whether a variable is numerical or categorical---and therefore whether you
> need to perform regression or classification---by taking two response variables X and Y from your
> data, and asking the question "is response variable X *more* than response variable Y?"
> If the variable is categorical, the question will make no sense ("is blue more than red?",
> or "is benign more than malignant?"). If the variable is numerical, it will make sense
> ("is 1.5 hours more than 2.25 hours?", or "is \$500,000 more than \$400,000?").
> Be careful when applying this heuristic, though: sometimes a categorical variables will be encoded as
> numbers in your data (e.g. "1" represents "benign", and "0" represents "malignant"). In these cases
> you have to ask the question about the *meaning* of the labels ("benign" and "malignant"), not their values ("1" and "0"). 

## Exploring a labelled data set

In this chapter and the next, we will study the Sacramento real estate data
set. This data set contains 932 real estate transactions in Sacramento,
California from a five day 
period [originally reported in the Sacramento Bee newspaper](https://support.spatialkey.com/spatialkey-sample-csv-data/).

```{r 07-load, message = FALSE}
library(tidyverse)
library(tidymodels)
library(gridExtra)

sacramento <- read_csv("data/sacramento.csv")
sacramento
```

We first need to formulate a precise question that
we want to answer. In this example, our question is predictive:
can we use the size of a house in the Sacramento, CA area to predict
its sale price? A rigorous, quantitative answer to this question might help
a realtor advise a client as to whether the price of a particular listing 
is fair, or perhaps how to set the price of a new listing.

The scientific question guides our initial exploration: the columns in the
data that we are interested in are `sqft` (house size, in livable square feet)
and `price` (house price, in US dollars (USD)).  The first step is to visualize
the data as a scatter plot where we place the predictor/explanatory variable
(house size) on the x-axis, and we place the target/response variable that we
want to predict (price) on the y-axis.

```{r 07-edaRegr, message = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage)"}
eda <- ggplot(sacramento, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format())
eda
```
 
The plot is shown in Figure \@ref(fig:07-edaRegr).
We can see that in Sacramento, CA, as the
size of a house increases, so does its sale price. Thus, we can reason that we
may be able to use the size of a not-yet-sold house (for which we don't know
the sale price) to predict its final sale price. Note that we do not suggest here
that a larger house size *causes* a higher sale price; just that house price
tends to increase with house size, and that we may be able to use the latter to 
predict the former.

## K-nearest neighbours regression

Much like in the case of classification, we can use a K-nearest
neighbours-based approach in regression to make predictions. Let's take a small
sample of the data above and walk through how K-nearest neighbours (knn) works
in a regression context before we dive in to creating our model and assessing
how well it predicts house price. This subsample is taken to allow us to
illustrate the mechanics of K-NN regression with a few data points; later in
this chapter we will use all the data.

To take a small random sample of size 30, we'll use the function `sample_n`.
This function takes two arguments:

1. `tbl` (a data frame-like object to sample from)
2. `size` (the number of observations/rows to be randomly selected/sampled)

```{r 07-sacramento}
set.seed(1)
small_sacramento <- sample_n(sacramento, size = 30)
```

Next let's say we come across a  2,000 square-foot house in Sacramento we are
interested in purchasing, with an advertised list price of \$350,000. Should we
offer to pay the asking price for this house, or is it overpriced and we should
offer less? Absent any other information, we can get a sense for a good answer
to this question by using the data we have to predict the sale price given the
sale prices we have already observed. But in Figure \ref@(fig:07-small-eda-regr),
you can see that we have no
observations of a house of size *exactly* 2000 square feet. How can we predict
the price? 

```{r 07-small-eda-regr, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with vertical line indicating 2000 square feet on x-axis"}
small_plot <- ggplot(small_sacramento, aes(x = sqft, y = price)) +
  geom_point() +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_vline(xintercept = 2000, linetype = "dotted")
small_plot
```

We will employ the same intuition from the classification chapter, and use the
neighbouring points to the new point of interest to suggest/predict what its
price should be. For the example above, we find and label the 5 nearest
neighbours to our observation of a house that is 2000 square feet.

```{r 07-find-k3}
nearest_neighbours <- small_sacramento %>%
  mutate(diff = abs(2000 - sqft)) %>%
  arrange(diff) %>%
  slice(1:5) #subset the first 5 rows
nearest_neighbours
```

```{r 07-knn3-example, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with lines to 5 nearest neighbours"}
nearest_neighbours <- mutate(nearest_neighbours, twothou = rep(2000, 5))
nn_plot <- small_plot +
  geom_segment(
    data = nearest_neighbours,
    aes(x = twothou, xend = sqft, y = price, yend = price), color = "orange"
  )
nn_plot
```

Figure \@ref(fig:07-knn3-example) illustrates the difference between the house sizes
of the 5 nearest neighbours (in terms of house size) to our new
2,000 square-foot house of interest. Now that we have obtained these nearest neighbours, 
we can use their values to predict a
selling price for the new home.  Specifically, we can take the mean (or
average) of these 5 values as our predicted value, as illustrated by
the red point in Figure \@ref(fig:07-predictedViz-knn).

```{r 07-predicted-knn}
prediction <- nearest_neighbours %>%
  summarise(predicted = mean(price))
prediction
```

```{r 07-predictedViz-knn, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with predicted price for a 2000 square-foot house based on 5 nearest neighbours represented as a red dot"}
nn_plot +
  geom_point(aes(x = 2000, y = prediction[[1]]), color = "red", size = 2.5)
```

Our predicted price is \$`r format(round(prediction[[1]]), scientific = FALSE)`
(shown as a red point above), which is much less than \$350,000; perhaps we
might want to offer less than the list price at which the house is advertised.
But this is only the very beginning of the story. We still have all the same
unanswered questions here with K-NN regression that we had with K-NN
classification: which $K$ do we choose, and is our model any good at making
predictions? In the next few sections, we will address these questions in the
context of K-NN regression.

## Training, evaluating, and tuning the model

As usual, we must start by putting some test data away in a lock box that we
will come back to only after we choose our final model. Let's take care of that
now. Note that for the remainder of the chapter we'll be working with the
entire Sacramento data set, as opposed to the smaller sample of 30 points above.

```{r 07-test-train-split}
set.seed(1234)
sacramento_split <- initial_split(sacramento, prop = 0.6, strata = price)
sacramento_train <- training(sacramento_split)
sacramento_test <- testing(sacramento_split)
```

Next, we'll use cross-validation to choose $K$. In K-NN classification, we used
accuracy to see how well our predictions matched the true labels. Here in the
context of K-NN regression we will use root mean square prediction error
(RMSPE) instead. The mathematical formula for calculating RMSPE is: 

$$\text{RMSPE} = \sqrt{\frac{1}{n}\sum\limits_{i=1}^{n}(y_i - \hat{y}_i)^2}$$

Where:

- $n$ is the number of observations
- $y_i$ is the observed value for the $i^\text{th}$ observation
- $\hat{y}_i$ is the forcasted/predicted value for the $i^\text{th}$ observation

A key feature of the formula for RMPSE is the squared difference between the observed
target/response variable value, $y$, and the prediction target/response
variable value, $\hat{y}_i$, for each observation (from 1 to $i$).
If the predictions are very close to the true values, then
RMSPE will be small. If, on the other-hand, the predictions are very
different to the true values, then RMSPE will be quite large. When we
use cross validation, we will choose the $K$ that gives
us the smallest RMSPE.

> **RMSPE versus RMSE**
When using many code packages (`tidymodels` included), the evaluation output 
we will get to assess the prediction quality of
our K-NN regression models is labelled "RMSE", or "root mean squared
error". Why is this so, and why not just RMSPE? 
In statistics, we try to be very precise with our
language to indicate whether we are calculating the prediction error on the
training data (*in-sample* prediction) versus on the testing data 
(*out-of-sample* prediction). When predicting and evaluating prediction quality on the training data, we 
 say RMSE. By contrast, when predicting and evaluating prediction quality
on the testing or validation data, we say RMSPE. 
The equation for calculating RMSE and RMSPE is exactly the same; all that changes is whether the $y$s are
training or testing data. But many people just use RMSE for both, 
and rely on context to denote which data the root mean squared error is being calculated on.

Now that we know how we can assess how well our model predicts a numerical
value, let's use R to perform cross-validation and to choose the optimal $K$.
First, we will create a model specification for K-nearest neighbours regression,
as well as a recipe for preprocessing our data. Note that we use `set_mode("regression")`
now in the model specification to denote a regression problem, as opposed to the classification
problems from the previous chapters. Note also that we include standardization
in our preprocessing to build good habits, but since we only have one 
predictor it is technically not necessary; there is no risk of comparing two predictors
of different scales.

```{r 07-choose-k-knn}
sacr_recipe <- recipe(price ~ sqft, data = sacramento_train) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())

sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sacr_vfold <- vfold_cv(sacramento_train, v = 5, strata = price)

sacr_wkflw <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec)
sacr_wkflw
```
The major difference you can see in the above workflow compared to previous
chapters is that we are running regression rather than classification. The fact
that we use `set_mode("regression")` essentially
tells `tidymodels` that we need to use different metrics (RMSPE, not accuracy)
for tuning and evaluation. You can see this in the following code, which tunes
the model and returns the RMSPE for each number of neighbours. 

```{r 07-choose-k-knn-results}

gridvals <- tibble(neighbors = seq(1, 200))

sacr_results <- sacr_wkflw %>%
  tune_grid(resamples = sacr_vfold, grid = gridvals) %>%
  collect_metrics()

# show all the results
sacr_results
```
We take the *minimum* RMSPE to find the best setting for the number of neighbours:
```{r 07-choose-k-knn-results-2}
# show only the row of minimum RMSPE
sacr_min <- sacr_results %>%
  filter(.metric == "rmse") %>%
  filter(mean == min(mean))
sacr_min
```
Here we can see that the smallest RMSPE occurs when $K =$ `r sacr_min %>% pull(neighbors)`.

## Underfitting and overfitting
Similar to the setting of classification, by setting the number of neighbours
to be too small or too large, we cause the RMSPE to increase, as shown in
Figure \@ref(fig:07-choose-k-knn-plot).  What is happening here? 

```{r 07-choose-k-knn-plot, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Effect of the number of neighbours on the RMSPE"} 
sacr_results <- sacr_results %>%
  filter(.metric == "rmse")

sacr_tunek_plot <- ggplot(sacr_results, aes(x = neighbors, y = mean)) +
  geom_point() +
  geom_line() +
  labs(x = "Neighbours", y = "RMSPE")

sacr_tunek_plot
```

Figure \@ref(fig:07-howK) visualizes the effect of different settings of $K$ on the
regression model. Each plot shows the predicted values for house price from
our K-NN regression model for 6 different values for $K$: 1, 3, `r kmin`, 41, 250, and 450.
For each model, we predict a price for every possible home size across the range of home sizes we
observed in the data set (here 500 to 4250 square feet) and we plot the
predicted prices as a blue line.

```{r 07-howK, echo = FALSE, fig.height = 12, fig.width = 10,fig.cap = "Predicted values for house price (represented as a blue line) from K-NN regression models for six different values for $K$"}
kmin <- sacr_min %>% pull(neighbors)
gridvals <- c(1, 3, kmin, 41, 250, 450)

plots <- list()

for (i in 1:6) {
  sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = gridvals[[i]]) %>%
    set_engine("kknn") %>%
    set_mode("regression")

  sacr_wkflw <- workflow() %>%
    add_recipe(sacr_recipe) %>%
    add_model(sacr_spec)

  sacr_preds <- sacr_wkflw %>%
    fit(data = sacramento_train) %>%
    predict(sacramento_train) %>%
    bind_cols(sacramento_train)

  plots[[i]] <- ggplot(sacr_preds, aes(x = sqft, y = price)) +
    geom_point(alpha = 0.4) +
    xlab("House size (square footage)") +
    ylab("Price (USD)") +
    scale_y_continuous(labels = dollar_format()) +
    geom_line(data = sacr_preds, aes(x = sqft, y = .pred), color = "blue") +
    ggtitle(paste0("K = ", gridvals[[i]]))
}
grid.arrange(grobs = plots, ncol = 2)
```

Figure \@ref(fig:07-howK) shows that when $K$ = 1, the blue line runs perfectly
through almost all of our training observations. This happens because our
predicted values for a given region depend on just a single observation. 
Since the flexible blue line follows the training observations very closely, if
we were to change any one of the training observation data points, we would
change the flexible blue line quite a lot. This means that if we
were to collect another training data set from the Sacramento real estate
market it likely wouldn't match those observations as well. Recall from the classification
chapters that this behaviour is called *overfitting*; we use this same term 
in the context of regression. We also sometimes say that the predictions have
*high variance*, because they *vary* a lot when you change the data a little bit.

What about the plot in Figure \@ref(fig:07-howK) where $K$ is quite large, say, $K$ = 450? 
In this case  the blue line is extremely smooth, and almost
flat. This happens because our predicted values for a given x value (here home
size), depend on many many (450) neighbouring observations. 
In contrast to the $K=1$ example, 
the smooth, inflexible blue line does not follow the training observations very
closely, and changing a single training data point
will likely not change the blue line very much.
This means that although the blue line matches does not match the data we happen to
have in this particular training data set perfectly, if we were to collect
another training data set from the Sacramento real estate market it likely
would match those observations equally poorly. Recall from the classification
chapters that this behaviour is called *underfitting*; we again use this same
term in the context of regression. We also sometimes say that the predictions
have *low variance*, because they don't *vary* too much when you change the 
data a little bit.

Ideally, what we want is neither of the two situations discussed above. Instead,
we would like a model with low variance (so that it will transfer/generalize
well to other data sets, and isn't too dependent on the
observations that happen to be in the training set) **and** low error on the training data
(where the model does not completely ignore our training data). If we explore 
the other values for $K$, in particular $K$ = `r sacr_min %>% pull(neighbors)`
(as suggested by cross-validation),
we can see it has a lower training data error than our model with a very high $K$ (e.g., 450),
and thus the model/predicted values better match the actual observed values
than the high $K$ model. Additionally, it has lower variance than our model
with a very low $K$ (e.g., 1) and thus it should better transer/generalize to
other data sets compared to the low $K$ model. All of this is similar to how
the choice of $K$ affects K-NN classification, as discussed in the previous
chapter. 

## Evaluating on the test set

To assess how well our model might do at predicting on unseen data, we will
assess its RMSPE on the test data. To do this, we will first
re-train our K-NN regression model on the entire training data set, 
using $K =$ `r sacr_min %>% pull(neighbors)` neighbours. Then we will
use `predict` to make predictions on the test data, and use the `metrics`
function again to compute the summary of regression quality. Because
we specify that we are performing regression in `set_mode`, the `metrics`
function knows to output a quality summary related to regression, and not, say, classification.

```{r 07-predict}
set.seed(1234)
kmin <- sacr_min %>% pull(neighbors)
sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = kmin) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sacr_fit <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec) %>%
  fit(data = sacramento_train)

sacr_summary <- sacr_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred)

sacr_summary
```
Our final model's test error as assessed by RMSPE 
is `r format(round(sacr_summary %>% filter(.metric == 'rmse') %>% pull(.estimate)), scientific=FALSE)`. 
But what does this RMSPE score mean? When we calculated test set prediction accuracy in K-NN
classification, the highest possible value was 1 and the lowest possible value was 0. 
If we got a value close to 1, our model was "good;" and otherwise,
the model was "not good." What about RMSPE? Unfortunately there is no default scale 
for RMSPE. Instead,  it is measured in
the units of the target/response variable, and so it is a bit hard to
interpret. For now, let's consider this approach to thinking about RMSPE from
our testing data set: as long as its not significantly worse than the cross-validation
RMSPE of our best model, then we can say that we're not doing too much worse on
the test data than we did on the training data. So the model appears to be
generalizing well to a new data set it has never seen before. In future courses
on statistical/machine learning, you will learn more about how to interpret RMSPE
from testing data and other ways to assess models.

Finally, Figure \@ref(fig:07-predict-all) shows the predictions that our final model makes across
all house sizes we might encounter in the Sacramento area. You have already seen a few
plots like this in this chapter; but here we also provide the code that generated it
as a learning challenge.

```{r 07-predict-all, fig.height = 4, fig.width = 5, fig.cap = "Predicted values of house price (blue line) for the final K-NN regression model"}
set.seed(1234)
sacr_preds <- sacr_fit %>%
  predict(sacramento_train) %>%
  bind_cols(sacramento_train)

plot_final <- ggplot(sacr_preds, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sacr_preds, aes(x = sqft, y = .pred), color = "blue") +
  ggtitle(paste0("K = ", kmin))

plot_final
```

## Strengths and limitations of K-NN regression

As with K-NN classification (or any prediction algorithm for that manner), K-NN regression has both strengths and weaknesses. Some are listed here:

**Strengths:**

K-nearest neighbours regression:

1. is simple and easy to understand
2. requires few assumptions about what the data must look like 
3. works well with non-linear relationships (i.e., if the relationship is not a straight line)

**Weaknesses:**

K-nearest neighbours regression:

1. becomes very slow as the training data gets larger
2. may not perform well with a large number of predictors
3. may not predict well beyond the range of values input in your training data

## Multivariate K-NN regression

As in K-NN classification, we can use multiple predictors in K-NN regression.
In this setting, we have the same concerns regarding the scale of the predictors. Once again, 
 predictions are made by identifying the $K$
observations that are nearest to the new point we want to predict; any
variables that are on a large scale will have a much larger effect than
variables on a small scale. But since the `recipe` we built above scales and centers
all predictor variables, this is handled for us.

Note that we also have the same concern regarding the selection of predictors
in K-NN regression as in K-NN classification: more predictors is **not** always
better, and the choice of which predictors to use has a potentially large influence
on the quality of predictions. Fortunately, we can use the best subset selection 
algorithm from the classification chapter in K-NN regression as well, assuming
a small number (at most 10 or so) of total predictors.

We will now demonstrate a multivariate K-NN regression analysis of the 
Sacramento real estate data using `tidymodels`. This time we will use
house size (measured in square feet) as well as number of bedrooms as our
predictors, and continue to use house sale price as our outcome/target variable
that we are trying to predict.

It is always a good practice to do exploratory data analysis, such as
visualizing the data, before we start modeling the data. Thus the first thing
we will do is use ggpairs (from the `GGally` package) to plot all the variables
we are interested in using in our analyses.

```{r 07-ggpairs, fig.height = 5, fig.width = 6, fig.cap = "Scatterplots of each pair of variables (price, house size, and the number of bedrooms) are displayed in the figure's bottom left corner. Correlation coefficients are shown in the top right corner. The distributions of price, house size and number of bedrooms are each shown along the diagonal."}
library(GGally)
plot_pairs <- sacramento %>%
  select(price, sqft, beds) %>%
  ggpairs()
plot_pairs
```

Figure \@ref(fig:07-ggpairs) shows that generally, as both house size and
number of bedrooms increase, so does price. Does adding the number of bedrooms
to our model improve our ability to predict house price? To answer that
question, we will have to estimate the test error for a K-NN regression
model using house size and number of bedrooms, and then we can compare it to
the test error for the model we previously came up with that only used house
size to see if it is smaller (decreased test error indicates increased
prediction quality). Let's do that now!

First we'll build a new model specification and recipe for the analysis. Note that
we use the formula `price ~ sqft + beds` to denote that we have two predictors,
and set `neighbors = tune()` to tell `tidymodels` to tune the number of neighbours for us.
```{r 07-mult-setup}
sacr_recipe <- recipe(price ~ sqft + beds, data = sacramento_train) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())

sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("regression")
```
Next, we'll use 5-fold cross-validation to choose the number of neighbours via the minimum RMSPE:
```{r 07-mult-cv}
gridvals <- tibble(neighbors = seq(1, 200))
sacr_k <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec) %>%
  tune_grid(sacr_vfold, grid = gridvals) %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  filter(mean == min(mean)) %>%
  pull(neighbors)
sacr_k
```
Here we see that the smallest RMSPE occurs when $K =$ `r sacr_k`.
Now that we have chosen $K$, we need to re-train the model on the entire
training data set, and after that we can use that model to predict
on the test data to get our test error. 
```{r 07-re-train}
sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = sacr_k) %>%
  set_engine("kknn") %>%
  set_mode("regression")

knn_mult_fit <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec) %>%
  fit(data = sacramento_train)

knn_mult_preds <- knn_mult_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test)

knn_mult_mets <- metrics(knn_mult_preds, truth = price, estimate = .pred)
knn_mult_mets
```
This time when we performed K-NN regression on the same data set, but also
included number of bedrooms as a predictor, we obtained a RMSPE test error 
of `r format(round(knn_mult_mets %>% filter(.metric == 'rmse') %>% pull(.estimate)), scientific=FALSE)`.
This compares to a RMSPE test error 
of  `r format(round(sacr_summary %>% filter(.metric == 'rmse') %>% pull(.estimate)), scientific=FALSE)`
when we used only house size as the
single predictor. Thus in this case, we did not improve the model 
by a large amount by adding this additional predictor.

Figure \@ref(fig:07-knn-mult-viz) visualizes the model's predictions overlaid on top of the data. This 
time the predictions are a surface in 3-D space, instead of a line in 2-D space, as we have 2
predictors instead of 1.  
```{r 07-knn-mult-viz, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "K-NN regression model’s predictions represented as a surface in 3-D space overlaid on top of the data using three predictors (price, house size, and the number of bedrooms)"}
library(plotly)

xvals <- seq(from = min(sacramento_train$sqft), to = max(sacramento_train$sqft), length = 50)
yvals <- seq(from = min(sacramento_train$beds), to = max(sacramento_train$beds), length = 50)

zvals <- knn_mult_fit %>%
  predict(crossing(xvals, yvals) %>% mutate(sqft = xvals, beds = yvals)) %>%
  pull(.pred)

zvalsm <- matrix(zvals, nrow = length(xvals))

plot_ly() %>%
  add_markers(
    data = sacramento_train,
    x = ~sqft,
    y = ~beds,
    z = ~price,
    marker = list(size = 5, opacity = 0.4, color = "red")
  ) %>%
  layout(scene = list(
    xaxis = list(title = "House size (square feet)"),
    zaxis = list(title = "Price (USD)"),
    yaxis = list(title = "Number of bedrooms")
  )) %>%
  add_surface(
    x = ~xvals,
    y = ~yvals,
    z = ~zvalsm,
    colorbar = list(title = "Price (USD)")
  )
```

We can see that the predictions in this case, where we have 2 predictors, form
a surface instead of a line. Because the newly added predictor (number of bedrooms) is 
related to price (as price changes, so does number of bedrooms)
and is not totally determined by house size (our other predictor),
we get additional and useful information for making our
predictions. For example, in this model we would predict that the cost of a
house with a size of 2,500 square feet generally increases slightly as the number
of bedrooms increases. Without having the additional predictor of number of
bedrooms, we would predict the same price for these two houses.
