# Regression II: linear regression {#regression2}

## Overview 
Up until this point, we have solved all of our predictive problems---both classification
and regression---using K-nearest neighbours-based approaches. But while K-NN methods are 
simple and flexible, in the context of regression, there is another method (generally more popular in
most scientific disciplines) known as *linear regression*. This chapter provides an introduction
to the basic concept of linear regression, shows how to use `tidymodels` to perform linear regression in R,
and characterizes its strengths and weaknesses compared to K-NN regression. The focus is, as usual,
on the case where there is a single predictor and single response variable of interest; but the chapter
concludes with an example using *multivariate linear regression* when there is more than one
predictor.

## Chapter learning objectives 
By the end of the chapter, students will be able to:

* Perform linear regression in R using `tidymodels` and evaluate it on a test data set.
* Compare and contrast predictions obtained from K-nearest neighbour regression to those obtained using simple ordinary least squares regression from the same dataset.
* In R, overlay regression lines from `geom_smooth` on a single plot. 

## Simple linear regression

At the end of the previous chapter, we noted some limitations of K-NN regression.
While the method is simple and easy to understand, K-NN regression does not
predict well beyond the range of the predictors in the training data, and
the method gets significantly slower as the training dataset grows.
Fortunately, there is an alternative to K-NN regression---*linear regression*---that addresses
both of these limitations. Linear regression is also much more commonly used in practice, especially
in scientific applications. In this first part of the chapter, we will focus on *simple* linear regression,
which involves only one predictor variable and one response variable; later on, we will consider
*multivariate* linear regression, which involves multiple predictor variables.
Like K-NN regression, simple linear regression involves
predicting a numerical response variable (like race time, house price, or height);
but *how* it makes those predictions for a new observation is quite different from K-NN regression.
 Instead of looking at the K nearest neighbours and averaging
over their values for a prediction, in simple linear regression, all the
training data points are used to create a straight line of best fit, and then
the line is used to "look up" the prediction. 

> **Note:** although we did not cover it explicitly in the classification chapters,
> there is another popular method for classification called *logistic regression* (ignore 
> the word *regression* there...). After reading this chapter, if you want a technique like
> linear regression but for classification, consider looking up logistic regression; 
> logistic regression and K-NN classification have a very similar advantage/disadvantage comparison
> as linear regression and K-NN regression. See the "Additional Resources" section at the end of
> the classification chapters to learn more about logistic regression.

Let's return to the Sacramento housing data from the previous chapter to learn how to apply linear regression and compare it to K-NN regression. For now, we will consider 
a smaller version of the housing data to help make our visualizations clear.
Recall our predictive question: can we use the size of a house in the Sacramento, CA area to predict
its sale price? In particular, recall that we have come across a new 2,000-square foot house we are interested
in purchasing with an advertised list price of
\$350,000. Should we offer the list price, or is that over/undervalued?
To answer this question using simple linear regression, we use the data we have
to draw the straight line of best fit through our existing data points:

```{r 08-lin-reg1, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with line of best fit for subset of the Sacramento housing data set"}
library(tidyverse)
library(gridExtra)
library(caret)

data(Sacramento)
set.seed(1234)
small_sacramento <- sample_n(Sacramento, size = 30)

small_plot <- ggplot(small_sacramento, aes(x = sqft, y = price)) +
  geom_point() +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE)
small_plot
```

The equation for the straight line is: 

$$\text{house price} = \beta_0 + \beta_1 \cdot (\text{house size}),$$
where

- $\beta_0$ is the vertical intercept of the line (the value where the line cuts the vertical axis)
- $\beta_1$ is the slope of the line

Therefore using the data to find the line of best fit is equivalent to finding coefficients 
$\beta_0$ and $\beta_1$ that *parametrize* (correspond to) the line of best fit.
Once we have the coefficients, we can use the equation above to evaluate the predicted price given the value we
have for the predictor/explanatory variable&mdash;here 2,000 square feet. 

```{r 08-lin-reg2, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with line of best fit and predicted price for a 2000 square foot home represented as a red dot"}
small_model <- lm(price ~ sqft, data = small_sacramento)
prediction <- predict(small_model, data.frame(sqft = 2000))

small_plot +
  geom_vline(xintercept = 2000, linetype = "dotted") +
  geom_point(aes(x = 2000, y = prediction[[1]], color = "red", size = 2.5)) +
  theme(legend.position = "none")

print(prediction[[1]])
```

By using simple linear regression on this small data set to predict the sale price
for a 2,000 square foot house, we get a predicted value of 
\$`r format(round(prediction[[1]]), scientific = FALSE)`. But wait a minute...how
exactly does simple linear regression choose the line of best fit? Many
different lines could be drawn through the data points. We show some examples
below:

```{r 08-several-lines, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5,  fig.cap = "Scatter plot of price (USD) versus house size (square footage) with many possible lines that could be drawn through the data points"}

small_plot +
  geom_abline(intercept = -64542.23, slope = 190, color = "green") +
  geom_abline(intercept = -6900, slope = 175, color = "purple") +
  geom_abline(intercept = -64542.23, slope = 160, color = "red")
```

Simple linear regression chooses the straight line of best fit by choosing
the line that minimizes the **average** vertical distance between itself and
each of the observed data points. From the lines shown above, that is the blue
line. What exactly do we mean by the vertical distance between the predicted
values (which fall along the line of best fit) and the observed data points?
We illustrate these distances in the plot below with a red line:

```{r 08-verticalDistToMin,  echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with the vertical distances between the predicted values and the observed data points"}
small_sacramento <- small_sacramento %>%
  mutate(predicted = predict(small_model))
small_plot +
  geom_segment(data = small_sacramento, aes(xend = sqft, yend = predicted), colour = "red")
```

To assess the predictive accuracy of a simple linear regression model,
we use RMSPE&mdash;the same measure of predictive performance we used with K-NN regression.

## Linear regression in R

We can perform simple linear regression in R using `tidymodels` in a
very similar manner to how we performed K-NN regression. 
To do this, instead of creating a `nearest_neighbor` model specification with
the `kknn` engine, we use a `linear_reg` model specification
with the `lm` engine. Another difference is that we do not need to choose $K$ in the
context of linear regression, and so we do not need to perform cross validation.
Below we illustrate how we can use the usual `tidymodels` workflow to predict house sale
price given house size using a simple linear regression approach using the full
Sacramento real estate data set.

> An additional difference that you will notice below is that we do not standardize
> (i.e., scale and center) our predictors. In K-nearest neighbours models, recall that 
> the model fit changes depending on whether we standardize first or not. In linear regression,
> standardization does not affect the fit (it *does* affect the coefficients in the equation, though!).
> So you can standardize if you want&mdash;it won't hurt anything&mdash;but if you leave the
> predictors in their original form, the best fit coefficients are usually easier to interpret afterward.

As usual, we start by putting some test data away in a lock box that we
can come back to after we choose our final model. Let's take care of that now.

```{r 08-test-train-split}
set.seed(1234)
sacramento_split <- initial_split(sacramento, prop = 0.6, strata = price)
sacramento_train <- training(sacramento_split)
sacramento_test <- testing(sacramento_split)
```

Now that we have our training data, we will create the model specification
and recipe, and fit our simple linear regression model:
```{r 08-fitLM, fig.height = 4, fig.width = 5}
lm_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lm_recipe <- recipe(price ~ sqft, data = sacramento_train)

lm_fit <- workflow() %>%
  add_recipe(lm_recipe) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
lm_fit
```
Our coefficients are 
(intercept) $\beta_0=$ `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), scientific=FALSE)`
and (slope) $\beta_1=$ `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`.
This means that the equation of the line of best fit is
$$\text{house price} = `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), scientific=FALSE)` + `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`\cdot (\text{house size}),$$
and that the model predicts that houses 
start at \$`r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), scientific=FALSE)` for 0 square feet, and that
every extra square foot increases the cost of the house by \$`r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`. Finally, we predict on the test data set to assess how well our model does:

```{r 08-assessFinal}
lm_test_results <- lm_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred)
lm_test_results
```

Our final model's test error as assessed by RMSPE
is `r format(round(lm_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate)), scientific=FALSE)`. 
Remember that this is in units of the target/response variable, and here that
is US Dollars (USD). Does this mean our model is "good" at predicting house
sale price based off of the predictor of home size? Again answering this is
tricky to answer and requires to use domain knowledge and think about the
application you are using the prediction for. 

To visualize the simple linear regression model, we can plot the predicted house
price across all possible house sizes we might encounter superimposed on a scatter
plot of the original housing price data. There is a plotting function in 
the `tidyverse`, `geom_smooth`, that
allows us to do this easily by adding a layer on our plot with the simple
linear regression predicted line of best fit. The default for this adds a
plausible range to this line that we are not interested in at this point, so to
avoid plotting it, we provide the argument `se = FALSE` in our call to
`geom_smooth`.

```{r 08-lm-predict-all, fig.height = 4, fig.width = 5, warning = FALSE, message = FALSE, fig.cap = "Scatter plot of price (USD) versus house size (square footage) with line of best fit for complete Sacramento housing data set"}

lm_plot_final <- ggplot(sacramento_train, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE)
lm_plot_final
```

We can extract the coefficients from our model by accessing the
fit object that is output by the `fit` function; we first have to extract
it from the workflow using the `pull_workflow_fit` function, and then apply
the `tidy` function to convert the result into a data frame:

```{r 08-lm-get-coeffs}
coeffs <- tidy(pull_workflow_fit(lm_fit))
coeffs
```

## Comparing simple linear and K-NN regression

Now that we have a general understanding of both simple linear and K-NN
regression, we can start to compare and contrast these methods as well as the
predictions made by them. To start, let's look at the visualization of the
simple linear regression model predictions for the Sacramento real estate data
(predicting price from house size) and the "best" K-NN regression model
obtained from the same problem:

```{r 08-compareRegression, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 10, fig.cap = "Comparison of simple linear regression and K-NN regression"}
set.seed(1234)
sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = 30) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sacr_wkflw <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec)

sacr_fit <- sacr_wkflw %>%
  fit(data = sacramento_train)

sacr_preds <- sacr_fit %>%
  predict(sacramento_train) %>%
  bind_cols(sacramento_train)

sacr_rmse <- sacr_preds %>%
  metrics(truth = price, estimate = .pred) %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round(2)

sacr_rmspe <- sacr_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred) %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round()


knn_plot_final <- ggplot(sacr_preds, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square footage)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sacr_preds, aes(x = sqft, y = .pred), color = "blue") +
  ggtitle("K-NN regression") +
  annotate("text", x = 3500, y = 100000, label = paste("RMSPE =", sacr_rmspe))

lm_rmspe <- lm_test_results %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round()

lm_plot_final <- lm_plot_final +
  annotate("text", x = 3500, y = 100000, label = paste("RMSPE =", lm_rmspe)) +
  ggtitle("linear regression")

grid.arrange(lm_plot_final, knn_plot_final, ncol = 2)
```

What differences do we observe from the visualization above? One obvious
difference is the shape of the blue lines. In simple linear regression we are
restricted to a straight line, whereas in K-NN regression our line is much more
flexible and can be quite wiggly. But there is a major interpretability advantage in limiting the
model to a straight line. A 
straight line can be defined by two numbers, the
vertical intercept and the slope. The intercept tells us what the prediction is when
all of the predictors are equal to 0; and the slope tells us what unit increase in the target/response
variable we predict given a unit increase in the predictor/explanatory
variable. K-NN regression, as simple as it is to implement and understand, has no such
interpretability from its wiggly line. 

There can however also be a disadvantage to using a simple linear regression
model in some cases, particularly when the relationship between the target and
the predictor is not linear, but instead some other shape (e.g. curved or oscillating). In 
these cases the prediction model from a simple linear regression
will underfit (have high bias), meaning that model/predicted values does not
match the actual observed values very well. Such a model would probably have a
quite high RMSE when assessing model goodness of fit on the training data and
a quite high RMPSE when assessing model prediction quality on a test data
set. On such a data set, K-NN regression may fare better. Additionally, there
are other types of regression you can learn about in future courses that may do
even better at predicting with such data.

How do these two models compare on the Sacramento house prices data set? On 
the visualizations above we also printed the RMPSE as calculated from 
predicting on the test data set that was not used to train/fit the models. The RMPSE for the simple linear
regression model is slightly lower than the RMPSE for the K-NN regression model.
Considering that the simple linear regression model is also more interpretable,
if we were comparing these in practice we would likely choose to use the simple
linear regression model.

Finally, note that the K-NN regression model becomes "flat"
at the left and right boundaries of the data, while the linear model
predicts a constant slope. Predicting outside the range of the observed
data is known as *extrapolation*; K-NN and linear models behave quite differently
when extrapolating. Depending on the application, the flat
or constant slope trend may make more sense. For example, if our housing
data were slightly different, the linear model may have actually predicted 
a *negative* price for a small houses (if the intercept $\beta_0$ was negative),
which obviously does not match reality. On the other hand, the trend of increasing
house size corresponding to increasing house price probably continues for large houses, 
so the "flat" extrapolation of K-NN likely does not match reality. 

## Multivariate linear regression

As in K-NN classification and K-NN regression, we can move beyond the simple
case of one response variable and only one predictor and perform multivariate
linear regression where we can have multiple predictors. In this case we fit a
plane to the data, as opposed to a straight line.

To do this, we follow a very similar approach to what we did for
K-NN regression; but recall that we do not need to use cross-validation to choose any parameters,
nor do we need to standardize (i.e., center and scale) the data for linear regression. 
We demonstrate how to do this below using the Sacramento real estate data with both house size
(measured in square feet) as well as number of bedrooms as our predictors, and
continue to use house sale price as our outcome/target variable that we are
trying to predict. We will start by changing the formula in the recipe to 
include both the `sqft` and `beds` variables as predictors:

```{r 08-lm-mult-test-train-split}
lm_recipe <- recipe(price ~ sqft + beds, data = sacramento_train)
```

Now we can build our workflow and fit the model:
```{r 08-fitlm}
lm_fit <- workflow() %>%
  add_recipe(lm_recipe) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
lm_fit
```

And finally, we predict on the test data set to assess how well our model does:

```{r 08-assessFinal-multi}
lm_mult_test_results <- lm_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred)
lm_mult_test_results
```

In the case of two predictors, our linear regression creates a *plane* of best fit, shown below:

```{r 08-3DlinReg, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Simple linear regression model’s predictions represented as a plane overlaid on top of the data using three predictors (price, house size, and the number of bedrooms)"}
library(plotly)
xvals <- seq(from = min(sacramento_train$sqft), to = max(sacramento_train$sqft), length = 50)
yvals <- seq(from = min(sacramento_train$beds), to = max(sacramento_train$beds), length = 50)

zvals <- lm_fit %>%
  predict(crossing(xvals, yvals) %>% mutate(sqft = xvals, beds = yvals)) %>%
  pull(.pred)

zvalsm <- matrix(zvals, nrow = length(xvals))

plot_ly() %>%
  add_markers(
    data = sacramento_train,
    x = ~sqft,
    y = ~beds,
    z = ~price,
    marker = list(size = 5, opacity = 0.4, color = "red")
  ) %>%
  layout(scene = list(
    xaxis = list(title = "House size (square feet)"),
    zaxis = list(title = "Price (USD)"),
    yaxis = list(title = "Number of bedrooms")
  )) %>%
  add_surface(
    x = ~xvals,
    y = ~yvals,
    z = ~zvalsm,
    colorbar = list(title = "Price (USD)")
  )
```
We see that the predictions from linear regression with two predictors form a
flat plane. This is the hallmark of linear regression, and differs from the 
wiggly, flexible surface we get from other methods such as K-NN regression. 
 As discussed this can be advantageous in one aspect, which is that for each
predictor, we can get slopes/intercept from linear regression, and thus describe the
plane mathematically. We can extract those slope values from our model object
as shown below:
```{r 08-lm-multi-get-coeffs}
coeffs <- tidy(pull_workflow_fit(lm_fit))
coeffs
```
And then use those slopes to write a mathematical equation to describe the prediction plane:

$$\text{house price} = \beta_0 + \beta_1\cdot(\text{house size}) + \beta_2\cdot(\text{number of bedrooms}),$$
where:

- $\beta_0$ is the vertical intercept of the hyperplane (the value where it cuts the vertical axis)
- $\beta_1$ is the slope for the first predictor (house size)
- $\beta_2$ is the slope for the second predictor (number of bedrooms)

Finally, we can fill in the values for $\beta_0$, $\beta_1$ and $\beta_2$ from the model output above
to create the equation of the plane of best fit to the data: 
```{r 08-lm-multi-get-coeffs-hidden, echo = FALSE}
icept <- format(round(coeffs %>% filter(term == "(Intercept)") %>% pull(estimate)), scientific = FALSE)
sqftc <- format(round(coeffs %>% filter(term == "sqft") %>% pull(estimate)), scientific = FALSE)
bedsc <- format(round(coeffs %>% filter(term == "beds") %>% pull(estimate)), scientific = FALSE)
```

$$\text{house price} = `r icept` + `r sqftc`\cdot (\text{house size})  `r bedsc` \cdot (\text{number of bedrooms})$$

This model is more interpretable than the multivariate K-NN
regression model; we can write a mathematical equation that explains how
each predictor is affecting the predictions. But as always, we should look at
the test error and ask whether linear regression is doing a better job of
predicting compared to K-NN regression in this multivariate regression case. To
do that we can use this linear regression model to predict on the test data to
get our test error.

```{r 08-get-RMSPE}
lm_mult_test_results
```

We get that the RMSPE for the multivariate linear regression model 
of `r format(lm_mult_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate), scientific = FALSE)`. This prediction error
is less than the prediction error for the multivariate K-NN regression model,
indicating that we should likely choose linear regression for predictions of
house price on this data set. But we should also ask if this more complex
model is doing a better job of predicting compared to our simple linear
regression model with only a single predictor (house size). Revisiting last
section, we see that our RMSPE for our simple linear regression model with
only a single predictor was 
`r format(lm_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate), scientific = FALSE)`, 
which is slightly more than that of our more complex model. Our model with two predictors
provided a slightly better fit on test data than our model with just one. 

But should we always end up choosing a model with more predictors than fewer? 
 The answer is no; you never know what model will be the best until you go through the
process of comparing their performance on held-out test data. Exploratory 
data analysis can give you some hints, but until you look
at the prediction errors to compare the models you don't really know.
Additionally, here we compare test errors purely for the purposes of teaching.
In practice, when  you want to compare several regression models with
differing numbers of predictor variables, you should use
cross-validation on the training set only; in this case choosing the model is part
of tuning, so you cannot use the test data. There are several well known and more advanced
methods to do this that are beyond the scope of this course, and they include
backward or forward selection, and L1 or L2 regularization (also known as Lasso
and ridge regression, respectively).

## The other side of regression

So far in this textbook we have used regression only in the context of
prediction. However, regression is also a powerful method to understand and/or
describe the relationship between a quantitative response variable and
one or more explanatory variables. Extending the case we have been working with
in this chapter (where we are interested in house price as the outcome/response
variable), we might also be interested in describing the
individual effects of house size and the number of bedrooms on house price,
quantifying how big each of these effects are, and assessing how accurately we
can estimate each of these effects. This side of regression is the topic of
many follow-on statistics courses and beyond the scope of this course.

## Additional resources
- Pages 59-71 of [Introduction to Statistical Learning](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Seventh%20Printing.pdf) with Applications in R by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani
- Pages 104 - 109 of [An Introduction to Statistical Learning with Applications in R](https://www-bcf.usc.edu/~gareth/ISL/ISLR%20Seventh%20Printing.pdf) by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani
- [The `caret` Package](https://topepo.github.io/caret/index.html)
- Chapters 6 - 11 of [Modern Dive](https://moderndive.com/) Statistical Inference via Data Science by Chester Ismay and Albert Y. Kim
