# Regression II: linear regression {#regression2}

## Overview 
Up to this point, we have solved all of our predictive problems---both classification
and regression---using K-nearest neighbours (KNN)-based approaches. In the context of regression, 
there is another method commonly used 
in scientific disciplines known as *linear regression*. This chapter provides an introduction
to the basic concept of linear regression, shows how to use `tidymodels` to perform linear regression in R,
and characterizes its strengths and weaknesses compared to KNN regression. The focus is, as usual,
on the case where there is a single predictor and single response variable of interest; but the chapter
concludes with an example using *multivariate linear regression* when there is more than one
predictor.

## Chapter learning objectives 
By the end of the chapter, students will be able to:

* Use R and `tidymodels` to fit a linear regression model on training data.
* Evaluate the linear regression model on test data.
* Compare and contrast predictions obtained from K-nearest neighbour regression to those obtained using linear regression from the same dataset.
* In R, overlay predictions from linear regression on a scatter plot of data using `geom_smooth`.

## Simple linear regression

At the end of the previous chapter, we noted some limitations of KNN regression.
While the method is simple and easy to understand, KNN regression does not
predict well beyond the range of the predictors in the training data, and
the method gets significantly slower as the training dataset grows.
Fortunately, there is an alternative to KNN regression---*linear regression*---that addresses
both of these limitations. Linear regression is also much more commonly used in practice, especially
in scientific applications, because it provides an interpretable mathematical equation that describes
the relationship between the predictor and response variables. In this first part of the chapter, we will focus on *simple* linear regression,
which involves only one predictor variable and one response variable; later on, we will consider
*multivariate* linear regression, which involves multiple predictor variables.
Like KNN regression, simple linear regression involves
predicting a numerical response variable (like race time, house price, or height);
but *how* it makes those predictions for a new observation is quite different from KNN regression.
 Instead of looking at the K nearest neighbours and averaging
over their values for a prediction, in simple linear regression, we create a 
straight line of best fit through the training data and then
"look up" the prediction using the line.

> **Note:** although we did not cover it in earlier chapters, there 
> is another popular method for classification called *logistic
> regression* (it is used for classification even though the name, somewhat confusingly,
> has the word "regression" in it). In logistic regression---similar to linear regression---you
> "fit" the model to the training data and then "look up" the prediction for each new observation.
> Logistic regression and KNN classification have an advantage/disadvantage comparison 
> similar to that of linear regression and KNN
> regression. It is useful to have a good understanding of linear regression before learning about
> logistic regression. After reading this chapter, see the "Additional Resources" section at the end of the
> classification chapters to learn more about logistic regression.

Let's return to the Sacramento housing data from the previous chapter to learn
how to apply linear regression and compare it to KNN regression. For now, we
will consider 
a smaller version of the housing data to help make our visualizations clear.
Recall our predictive question: can we use the size of a house in the Sacramento, CA area to predict
its sale price? In particular, recall that we have come across a new 2,000-square foot house we are interested
in purchasing with an advertised list price of
\$350,000. Should we offer the list price, or is that over/undervalued?
To answer this question using simple linear regression, we use the data we have
to draw the straight line of best fit through our existing data points.
The small subset of data as well as the line of best fit are shown
in Figure \@ref(fig:08-lin-reg1).

```{r 08-lin-reg1, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of sale price versus size with line of best fit for subset of the Sacramento housing data"}
library(tidyverse)
library(tidymodels)
library(scales)
library(gridExtra)
library(caret)

sacramento <- read_csv("data/sacramento.csv")

set.seed(2)
small_sacramento <- sample_n(sacramento, size = 30)

small_plot <- ggplot(small_sacramento, aes(x = sqft, y = price)) +
  geom_point() +
  xlab("House size (square feet)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE)
small_plot
```

The equation for the straight line is: 

$$\text{house sale price} = \beta_0 + \beta_1 \cdot (\text{house size}),$$
where

- $\beta_0$ is the *vertical intercept* of the line (the price when house size is 0)
- $\beta_1$ is the *slope* of the line (how quickly the price increases as you increase house size)

Therefore using the data to find the line of best fit is equivalent to finding coefficients 
$\beta_0$ and $\beta_1$ that *parametrize* (correspond to) the line of best fit.
Now of course, in this particular problem, the idea of a 0-square-foot house is a bit silly;
but you can think of $\beta_0$ here as the "base price," and 
$\beta_1$ as the increase in price for each square foot of space.
Let's push this thought even further: what would happen in the equation for the line if you 
tried to evaluate the price of a house with size 6 *million* square feet?
Or what about *negative* 2,000 square feet? As it turns out, nothing in the formula breaks; linear
regression will happily make predictions for crazy predictor values if you ask it to. But even though
you *can* make these wild predictions, you shouldn't. You should only make predictions roughly within
the range of your original data, and perhaps a bit beyond it only if it makes sense. For example,
the data in Figure \@ref(fig:08-lin-reg1) only reaches around 800 square feet on the low end, but 
it would probably be reasonable to use the linear regression model to make a prediction at 600 square feet, say.

Back to the example! Once we have the coefficients $\beta_0$ and $\beta_1$, we can use the equation
above to evaluate the predicted sale price given the value we have for the
predictor variable&mdash;here 2,000 square feet. Figure
\@ref(fig:08-lin-reg2) demonstrates this process.

```{r 08-lin-reg2, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of sale price versus size with line of best fit and a red dot at the predicted sale price for a 2000 square foot home"}
small_model <- lm(price ~ sqft, data = small_sacramento)
prediction <- predict(small_model, data.frame(sqft = 2000))

small_plot +
  geom_vline(xintercept = 2000, linetype = "dotted") +
  geom_point(aes(x = 2000, y = prediction[[1]], color = "red", size = 2.5)) +
  annotate("text", x = 2350, y = prediction[[1]]-30000, label=paste("$", format(round(prediction[[1]]), big.mark=",", nsmall=0, scientific = FALSE), sep=""))+
  theme(legend.position = "none")
```

By using simple linear regression on this small data set to predict the sale price
for a 2,000 square foot house, we get a predicted value of 
\$`r format(round(prediction[[1]]), big.mark=",", nsmall=0, scientific = FALSE)`. But wait a minute...how
exactly does simple linear regression choose the line of best fit? Many
different lines could be drawn through the data points. 
Some plausible examples are shown in Figure \@ref(fig:08-several-lines).

```{r 08-several-lines, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5,  fig.cap = "Scatter plot of sale price versus size with many possible lines that could be drawn through the data points"}

small_plot +
  geom_abline(intercept = -64542.23, slope = 190, color = "green") +
  geom_abline(intercept = -6900, slope = 175, color = "purple") +
  geom_abline(intercept = -64542.23, slope = 160, color = "red")
```

Simple linear regression chooses the straight line of best fit by choosing
the line that minimizes the **average squared vertical distance** between itself and
each of the observed data points in the training data. Figure \@ref(fig:08-verticalDistToMin) illustrates 
these vertical distances as red lines. Finally, to assess the predictive 
accuracy of a simple linear regression model,
we use RMSPE&mdash;the same measure of predictive performance we used with KNN regression.

```{r 08-verticalDistToMin,  echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Scatter plot of sale price versus size with red lines denoting the vertical distances between the predicted values and the observed data points"}
small_sacramento <- small_sacramento %>%
  mutate(predicted = predict(small_model))
small_plot +
  geom_segment(data = small_sacramento, aes(xend = sqft, yend = predicted), colour = "red")
```


## Linear regression in R

We can perform simple linear regression in R using `tidymodels` in a
very similar manner to how we performed KNN regression. 
To do this, instead of creating a `nearest_neighbor` model specification with
the `kknn` engine, we use a `linear_reg` model specification
with the `lm` engine. Another difference is that we do not need to choose $K$ in the
context of linear regression, and so we do not need to perform cross validation.
Below we illustrate how we can use the usual `tidymodels` workflow to predict house sale
price given house size using a simple linear regression approach using the full
Sacramento real estate data set.

> An additional difference that you will notice below is that we do not standardize
> (i.e., scale and center) our predictors. In K-nearest neighbours models, recall that 
> the model fit changes depending on whether we standardize first or not. In linear regression,
> standardization does not affect the fit (it *does* affect the coefficients in the equation, though!).
> So you can standardize if you want&mdash;it won't hurt anything&mdash;but if you leave the
> predictors in their original form, the best fit coefficients are usually easier to interpret afterward.

As usual, we start by putting some test data away in a lock box that we
can come back to after we choose our final model. Let's take care of that now.

```{r 08-test-train-split}
set.seed(1234)
sacramento_split <- initial_split(sacramento, prop = 0.6, strata = price)
sacramento_train <- training(sacramento_split)
sacramento_test <- testing(sacramento_split)
```

Now that we have our training data, we will create the model specification
and recipe, and fit our simple linear regression model:
```{r 08-fitLM, fig.height = 4, fig.width = 5}
lm_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lm_recipe <- recipe(price ~ sqft, data = sacramento_train)

lm_fit <- workflow() %>%
  add_recipe(lm_recipe) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
lm_fit
```
Our coefficients are 
(intercept) $\beta_0=$ `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), scientific=FALSE)`
and (slope) $\beta_1=$ `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`.
This means that the equation of the line of best fit is
$$\text{house sale price} = `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), scientific=FALSE)` + `r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`\cdot (\text{house size}).$$
In other words, the model predicts that houses 
start at \$`r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[1]), big.mark=",", nsmall=0, scientific=FALSE)` for 0 square feet, and that
every extra square foot increases the cost of the house by \$`r format(round(pull(tidy(pull_workflow_fit(lm_fit)), estimate)[2]), scientific=FALSE)`. Finally, we predict on the test data set to assess how well our model does:

```{r 08-assessFinal}
lm_test_results <- lm_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred)
lm_test_results
```

Our final model's test error as assessed by RMSPE
is `r format(round(lm_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate)), big.mark=",", nsmall=0, scientific=FALSE)`. 
Remember that this is in units of the target/response variable, and here that
is US Dollars (USD). Does this mean our model is "good" at predicting house
sale price based off of the predictor of home size? Again, answering this is
tricky and requires knowledge of how you intend to use the prediction.

To visualize the simple linear regression model, we can plot the predicted house
sale price across all possible house sizes we might encounter superimposed on a scatter
plot of the original housing price data. There is a plotting function in 
the `tidyverse`, `geom_smooth`, that
allows us to add a layer on our plot with the simple
linear regression predicted line of best fit. By default `geom_smooth` adds some other information
to the plot that we are not interested in at this point; we provide the argument `se = FALSE` to
tell `geom_smooth` not to show that information. Figure \@ref(fig:08-lm-predict-all) displays the result.

```{r 08-lm-predict-all, fig.height = 4, fig.width = 5, warning = FALSE, message = FALSE, fig.cap = "Scatter plot of sale price versus size with line of best fit for the full Sacramento housing data"}

lm_plot_final <- ggplot(sacramento_train, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square feet)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE)
lm_plot_final
```

We can extract the coefficients from our model by accessing the
fit object that is output by the `fit` function; we first have to extract
it from the workflow using the `pull_workflow_fit` function, and then apply
the `tidy` function to convert the result into a data frame:

```{r 08-lm-get-coeffs}
coeffs <- lm_fit %>%
             pull_workflow_fit() %>%
             tidy()
coeffs
```

## Comparing simple linear and KNN regression

Now that we have a general understanding of both simple linear and KNN
regression, we can start to compare and contrast these methods as well as the
predictions made by them. To start, let's look at the visualization of the
simple linear regression model predictions for the Sacramento real estate data
(predicting price from house size) and the "best" KNN regression model
obtained from the same problem, shown in Figure \@ref(fig:08-compareRegression).

```{r 08-compareRegression, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 10, fig.cap = "Comparison of simple linear regression and KNN regression"}
set.seed(1234)
sacr_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = 30) %>%
  set_engine("kknn") %>%
  set_mode("regression")


sacr_recipe <- recipe(price ~ sqft, data = sacramento_train) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())

sacr_wkflw <- workflow() %>%
  add_recipe(sacr_recipe) %>%
  add_model(sacr_spec)

sacr_fit <- sacr_wkflw %>%
  fit(data = sacramento_train)

sacr_preds <- sacr_fit %>%
  predict(sacramento_train) %>%
  bind_cols(sacramento_train)

sacr_rmse <- sacr_preds %>%
  metrics(truth = price, estimate = .pred) %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round(2)

sacr_rmspe <- sacr_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred) %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round()


knn_plot_final <- ggplot(sacr_preds, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("House size (square feet)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sacr_preds, aes(x = sqft, y = .pred), color = "blue") +
  ggtitle("KNN regression") +
  annotate("text", x = 3500, y = 100000, label = paste("RMSPE =", sacr_rmspe))

lm_rmspe <- lm_test_results %>%
  filter(.metric == "rmse") %>%
  pull(.estimate) %>%
  round()

lm_plot_final <- lm_plot_final +
  annotate("text", x = 3500, y = 100000, label = paste("RMSPE =", lm_rmspe)) +
  ggtitle("linear regression")

grid.arrange(lm_plot_final, knn_plot_final, ncol = 2)
```

What differences do we observe in Figure \@ref(fig:08-compareRegression)? One obvious
difference is the shape of the blue lines. In simple linear regression we are
restricted to a straight line, whereas in KNN regression our line is much more
flexible and can be quite wiggly. But there is a major interpretability advantage in limiting the
model to a straight line. A 
straight line can be defined by two numbers, the
vertical intercept and the slope. The intercept tells us what the prediction is when
all of the predictors are equal to 0; and the slope tells us what unit increase in the target/response
variable we predict given a unit increase in the predictor
variable. KNN regression, as simple as it is to implement and understand, has no such
interpretability from its wiggly line. 

There can however also be a disadvantage to using a simple linear regression
model in some cases, particularly when the relationship between the target and
the predictor is not linear, but instead some other shape (e.g. curved or oscillating). In 
these cases the prediction model from a simple linear regression
will underfit (have high bias), meaning that model/predicted values does not
match the actual observed values very well. Such a model would probably have a
quite high RMSE when assessing model goodness of fit on the training data and
a quite high RMPSE when assessing model prediction quality on a test data
set. On such a data set, KNN regression may fare better. Additionally, there
are other types of regression you can learn about in future courses that may do
even better at predicting with such data.

How do these two models compare on the Sacramento house prices data set? In
Figure \@ref(fig:08-compareRegression), we also printed the RMPSE as calculated from 
predicting on the test data set that was not used to train/fit the models. The RMPSE for the simple linear
regression model is slightly lower than the RMPSE for the KNN regression model.
Considering that the simple linear regression model is also more interpretable,
if we were comparing these in practice we would likely choose to use the simple
linear regression model.

Finally, note that the KNN regression model becomes "flat"
at the left and right boundaries of the data, while the linear model
predicts a constant slope. Predicting outside the range of the observed
data is known as *extrapolation*; KNN and linear models behave quite differently
when extrapolating. Depending on the application, the flat
or constant slope trend may make more sense. For example, if our housing
data were slightly different, the linear model may have actually predicted 
a *negative* price for a small houses (if the intercept $\beta_0$ was negative),
which obviously does not match reality. On the other hand, the trend of increasing
house size corresponding to increasing house price probably continues for large houses, 
so the "flat" extrapolation of KNN likely does not match reality. 

## Multivariate linear regression

As in KNN classification and KNN regression, we can move beyond the simple
case of only one predictor to the case with multiple predictors,
known as *multivariate linear regression*. 
To do this, we follow a very similar approach to what we did for
KNN regression: we just add more predictors to the model formula in the 
recipe. But recall that we do not need to use cross-validation to choose any parameters,
nor do we need to standardize (i.e., center and scale) the data for linear regression. 
Note once again that we have the same concerns regarding multiple predictors
as in the settings of multivariate KNN regression and classification: more predictors is **not** always
better. But because the same predictor selection 
algorithm from the classification chapter extends to the setting of linear regression,
it will not be covered again in this chapter.

We will demonstrate multivariate linear regression using the Sacramento real estate 
data with both house size
(measured in square feet) as well as number of bedrooms as our predictors, and
continue to use house sale price as our response variable. We will start by 
changing the formula in the recipe to 
include both the `sqft` and `beds` variables as predictors:

```{r 08-lm-mult-test-train-split}
mlm_recipe <- recipe(price ~ sqft + beds, data = sacramento_train)
```

Now we can build our workflow and fit the model:
```{r 08-fitlm}
mlm_fit <- workflow() %>%
  add_recipe(mlm_recipe) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
mlm_fit
```

And finally, we make predictions on the test data set to assess the quality of our model:

```{r 08-assessFinal-multi}
lm_mult_test_results <- mlm_fit %>%
  predict(sacramento_test) %>%
  bind_cols(sacramento_test) %>%
  metrics(truth = price, estimate = .pred)
lm_mult_test_results
```
Our model's test error as assessed by RMSPE
is `r format(round(lm_mult_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate)), big.mark=",", nsmall=0, scientific=FALSE)`,
  

In the case of two predictors, our linear regression creates a *plane* of best fit, as
shown in Figure \@ref(fig:08-3DlinReg).

```{r 08-3DlinReg, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Linear regression plane of best fit overlaid on top of the data (using price, house size, and number of bedrooms as predictors)"}
library(plotly)
xvals <- seq(from = min(sacramento_train$sqft), to = max(sacramento_train$sqft), length = 50)
yvals <- seq(from = min(sacramento_train$beds), to = max(sacramento_train$beds), length = 50)

zvals <- mlm_fit %>%
  predict(crossing(xvals, yvals) %>% mutate(sqft = xvals, beds = yvals)) %>%
  pull(.pred)

zvalsm <- matrix(zvals, nrow = length(xvals))

plot_ly() %>%
  add_markers(
    data = sacramento_train,
    x = ~sqft,
    y = ~beds,
    z = ~price,
    marker = list(size = 5, opacity = 0.4, color = "red")
  ) %>%
  layout(scene = list(
    xaxis = list(title = "House size (square feet)"),
    zaxis = list(title = "Price (USD)"),
    yaxis = list(title = "Number of bedrooms")
  )) %>%
  add_surface(
    x = ~xvals,
    y = ~yvals,
    z = ~zvalsm,
    colorbar = list(title = "Price (USD)")
  )
```
We see that the predictions from linear regression with two predictors form a
flat plane. This is the hallmark of linear regression, and differs from the 
wiggly, flexible surface we get from other methods such as KNN regression. 
 As discussed this can be advantageous in one aspect, which is that for each
predictor, we can get slopes/intercept from linear regression, and thus describe the
plane mathematically. We can extract those slope values from our model object
as shown below:
```{r 08-lm-multi-get-coeffs}
mcoeffs <- mlm_fit %>%
             pull_workflow_fit() %>%
             tidy()
mcoeffs
```
And then use those slopes to write a mathematical equation to describe the prediction plane:

$$\text{house sale price} = \beta_0 + \beta_1\cdot(\text{house size}) + \beta_2\cdot(\text{number of bedrooms}),$$
where:

- $\beta_0$ is the *vertical intercept* of the hyperplane (the price when both house size and number of bedrooms are 0)
- $\beta_1$ is the *slope* for the first predictor (how quickly the price increases as you increase house size)
- $\beta_2$ is the *slope* for the second predictor (how quickly the price increases as you increase the number of bedrooms)

Finally, we can fill in the values for $\beta_0$, $\beta_1$ and $\beta_2$ from the model output above
to create the equation of the plane of best fit to the data: 
```{r 08-lm-multi-get-coeffs-hidden, echo = FALSE}
icept <- format(round(mcoeffs %>% filter(term == "(Intercept)") %>% pull(estimate)), scientific = FALSE)
sqftc <- format(round(mcoeffs %>% filter(term == "sqft") %>% pull(estimate)), scientific = FALSE)
bedsc <- format(round(mcoeffs %>% filter(term == "beds") %>% pull(estimate)), scientific = FALSE)
```

$$\text{house sale price} = `r icept` + `r sqftc`\cdot (\text{house size})  `r bedsc` \cdot (\text{number of bedrooms})$$

This model is more interpretable than the multivariate KNN
regression model; we can write a mathematical equation that explains how
each predictor is affecting the predictions. But as always, we should 
question how well multivariate linear regression is doing compared to
the other tools we have, such as simple linear regression 
and multivariate KNN regression. If this comparison is part of 
the model tuning process&mdash;for example, if we are trying
out many different sets of predictors for multivariate linear
and KNN regression&mdash;we must perform this comparison using
cross-validation on only our training data. But if we have already
decided on a small number (e.g., 2 or 3) of tuned candidate models and
we want to make a final comparison, we can do so by comparing the prediction
error of the methods on the test data.

```{r 08-get-RMSPE}
lm_mult_test_results
```

We obtain an RMSPE for the multivariate linear regression model 
of `r format(lm_mult_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate), big.mark=",", nsmall=0, scientific = FALSE)`. This prediction error
is less than the prediction error for the multivariate KNN regression model,
indicating that we should likely choose linear regression for predictions of
house sale price on this data set. Revisiting the simple linear regression model
with only a single predictor from earlier in this chapter, we see that the RMSPE for that model was 
`r format(lm_test_results %>% filter(.metric == 'rmse') %>% pull(.estimate), big.mark=",", nsmall=0, scientific = FALSE)`, 
which is slightly higher than that of our more complex model. Our model with two predictors
provided a slightly better fit on test data than our model with just one. 
As mentioned earlier, this is not always the case: sometimes including more
predictors can negatively impact the prediction performance on unseen 
test data.

## Multicollinearity and outliers

What can go wrong when performing (multivariate) linear regression? 
Two common issues to be mindful of are *outliers* and, in the multivarate case, *collinear predictors*.

### Outliers

Outliers are data points that do not follow the usual pattern of the rest of the data.
In the setting of linear regression, these are points that
 have a vertical distance to the line of best fit that is either much higher or much lower 
than you might expect based on the rest of the data. The problem with outliers is that
they can have *too much influence* on the line of best fit. In general, it is very difficult
to judge accurately which data are outliers without advanced techniques that are beyond
the scope of this book.

But to illustrate what can happen when you have outliers, Figure \@ref(fig:08-lm-outlier)
shows a small subset of the Sacramento housing data again, except we have added a *single* data point (highlighted
in red). This house is 5,000 square feet in size, and sold for only \$50,000. Unbeknownst to the
data analyst, this house was sold by a parent to their child for an absurdly low price. Of course
this is not representative of the real housing market values that the other data points follow; 
the data point is an *outlier*. In blue we plot the original line of best fit, and in red
we plot the new line of best fit including the outlier. You can see how different the red line
is from the blue line, which is entirely caused by that one extra outlier data point.

```{r 08-lm-outlier, fig.height = 4, fig.width = 5, warning = FALSE, echo = FALSE, fig.cap = "Scatter plot of a subset of the data, with outlier highlighted in red"}

sacramento_train_small <- sacramento_train %>% sample_n(100)

sacramento_outlier <- tibble(sqft = 5000, price = 50000)

lm_plot_outlier <- ggplot(sacramento_train_small, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  geom_point(data = sacramento_outlier, mapping = aes(x = sqft, y = price), color="red", size = 2.5) +
  xlab("House size (square feet)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE) + 
  geom_smooth(data = sacramento_train_small %>% add_row(sqft = 5000, price = 50000), method = "lm", se = FALSE, color = "red")
lm_plot_outlier
```

Fortunately, if you have enough data, the inclusion of one or two
outliers&mdash;as long as their values are not *too* wild&mdash;will
typically not have a large an effect on the line of best fit. Figure
\@ref(fig:08-lm-outlier-2) shows how that same outlier data point from earlier
influences the line of best fit when we are working with the entire original
Sacramento training data. You can see that with this larger data set, the line
changes much less when adding the outlier.
Nevertheless, it is still important when working with linear regression to critically
think about how much any individual data point is influencing the model.

```{r 08-lm-outlier-2, fig.height = 4, fig.width = 5, warning = FALSE, echo = FALSE, fig.cap = "Scatter plot of the full data, with outlier highlighted in red"}

sacramento_outlier <- tibble(sqft = 5000, price = 50000)

lm_plot_outlier_large <- ggplot(sacramento_train, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  geom_point(data = sacramento_outlier, mapping = aes(x = sqft, y = price), color="red", size = 2.5) +
  xlab("House size (square feet)") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE) + 
  geom_smooth(data = sacramento_train %>% add_row(sqft = 5000, price = 50000), method = "lm", se = FALSE, color = "red")
lm_plot_outlier_large
```

### Multicollinearity

The second, and much more subtle, issue can occur when performing multivariate linear regression.
In particular, if you include multiple predictors that are strongly linearly related to one another,
the coefficients that describe the plane of best fit can be very unstable. Consider
an extreme example using the Sacramento housing data where the house was measured twice
by two people.
Since the two people are each slightly inaccurate, the two measurements might not agree exactly,
but they are very strongly linearly related to each other, as shown in Figure \@ref(fig:08-lm-multicol).

```{r 08-lm-multicol, fig.height = 4, fig.width = 5, warning = FALSE, echo = FALSE, fig.cap = "Scatter plot of the with possible outlier highlighted in red"}

sacramento_train <- sacramento_train %>%
                          mutate(sqft1 = sqft + 100*sample(1000000, size=nrow(sacramento_train), replace=TRUE)/1000000) %>%
                          mutate(sqft2 = sqft + 100*sample(1000000, size=nrow(sacramento_train), replace=TRUE)/1000000) %>%
                          mutate(sqft3 = sqft + 100*sample(1000000, size=nrow(sacramento_train), replace=TRUE)/1000000) 

lm_plot_multicol_1 <- ggplot(sacramento_train) +
  geom_point(aes(x = sqft, y = sqft1), alpha=0.4)+
  xlab("House size (square feet)") +
  ylab("House size (square inches)") 
lm_plot_multicol_1
```

```{r 08-lm-multicol-2, echo = FALSE, warning = FALSE}
lm_recipe1 <- recipe(price ~ sqft + sqft1, data = sacramento_train)
lm_fit1 <- workflow() %>%
  add_recipe(lm_recipe1) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
coeffs <- tidy(pull_workflow_fit(lm_fit1))
icept1 <- format(round(coeffs %>% filter(term == "(Intercept)") %>% pull(estimate)), scientific = FALSE)
sqft1 <- format(round(coeffs %>% filter(term == "sqft") %>% pull(estimate)), scientific = FALSE)
sqft11 <- format(round(coeffs %>% filter(term == "sqft1") %>% pull(estimate)), scientific = FALSE)

lm_recipe2 <- recipe(price ~ sqft + sqft2, data = sacramento_train)
lm_fit2 <- workflow() %>%
  add_recipe(lm_recipe2) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
coeffs <- tidy(pull_workflow_fit(lm_fit2))
icept2 <- format(round(coeffs %>% filter(term == "(Intercept)") %>% pull(estimate)), scientific = FALSE)
sqft2 <- format(round(coeffs %>% filter(term == "sqft") %>% pull(estimate)), scientific = FALSE)
sqft22 <- format(round(coeffs %>% filter(term == "sqft2") %>% pull(estimate)), scientific = FALSE)

lm_recipe3 <- recipe(price ~ sqft + sqft3, data = sacramento_train)
lm_fit3 <- workflow() %>%
  add_recipe(lm_recipe3) %>%
  add_model(lm_spec) %>%
  fit(data = sacramento_train)
coeffs <- tidy(pull_workflow_fit(lm_fit3))
icept3 <- format(round(coeffs %>% filter(term == "(Intercept)") %>% pull(estimate)), scientific = FALSE)
sqft3 <- format(round(coeffs %>% filter(term == "sqft") %>% pull(estimate)), scientific = FALSE)
sqft33 <- format(round(coeffs %>% filter(term == "sqft3") %>% pull(estimate)), scientific = FALSE)

```

If we again fit the multivariate linear regression model on this data, then the plane of best fit
has regression coefficients that are very sensitive to the exact values in the data. For example,
if we change the data ever so slightly&mdash;e.g., by running cross validation, which splits
up the data randomly into different chunks&mdash;the coefficients vary by large amounts:

Best Fit 1: $\text{house sale price} = `r icept1` + `r sqft1`\cdot (\text{house size 1 (ft$^2$)}) + `r sqft11` \cdot (\text{house size 2 (ft$^2$)}).$

Best Fit 2: $\text{house sale price} = `r icept2` + `r sqft2`\cdot (\text{house size 1 (ft$^2$)}) + `r sqft22` \cdot (\text{house size 2 (ft$^2$)}).$

Best Fit 3: $\text{house sale price} = `r icept3` + `r sqft3`\cdot (\text{house size 1 (ft$^2$)}) + `r sqft33` \cdot (\text{house size 2 (ft$^2$)}).$

Therefore, when performing multivariate linear regression, it is important to avoid including very
linearly related predictors. 

## Designing new predictors

We were quite fortunate in our initial exploration to find a predictor variable (house size)
that seems to have a meaningful and nearly linear relationship with our response variable (sale price).
But what should we do if we cannot immediately find such a nice variable?
Well, sometimes it is just a fact that the variables in the data do not have enough of 
a relationship with the response variable to provide useful predictions. For example,
if the only available predictor was "the current house owner's favourite ice cream flavour",
we likely would have little hope of using that variable to predict the house's sale price
(barring any future remarkable scientific discoveries about the relationship between 
the housing market and homeowner ice cream preferences). In cases like these, 
the only option is to obtain measurements of more useful variables.

There are, however, a wide variety of cases where the predictor variables do have a 
meaningful relationship with the response variable, but that relationship does not fit
the assumptions of the regression method you have chosen. For example, a dataframe `df`
with two variables---`x` and `y`---with a nonlinear relationship between the two variables
will not be correctly captured by simple linear regression, as shown in Figure \@ref(fig:08-predictor-design).
<center>
```{r 08-predictor-design, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Example of a dataset with a nonlinear relationship between the predictor and the response"}
set.seed(1)
df = tibble(x = sample(10000, 100, replace=TRUE)/10000)
df <- df %>% mutate(y = x^3 + 0.2*sample(10000, 100, replace=TRUE)/10000 - 0.1)
curve_plt <- ggplot(df, aes(x = x, y = y)) +
  geom_point() +
  xlab("x") +
  ylab("y") +
  geom_smooth(method = "lm", se = FALSE)
curve_plt
```
</center>
Instead of trying to predict the response `y` using a linear regression on `x`,
we might have some scientific background about our problem to suggest that `y`
should be a cubic function of `x`. So before performing regression,
we might *create a new predictor variable* `z` using the `mutate` function:
```{r 08-predictor-design-3, message = FALSE, warning = FALSE}
df <- df %>%
        mutate(z = x^3)
```
Then we can perform linear regression for `y` using the predictor variable `z`,
as shown in Figure \@ref(fig:08-predictor-design-2).
Here you can see that the transformed predictor `z` helps the 
linear regression model make more accurate predictions. 
Note that none of the `y` response values have changed between Figures \@ref(fig:08-predictor-design)
and \@ref(fig:08-predictor-design-2); the only change is that the `x` values
have been replaced by `z` values.
<center>
```{r 08-predictor-design-2, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 4, fig.width = 5, fig.cap = "Relationship between the transformed predictor and the response"}
curve_plt2 <- ggplot(df, aes(x = z, y = y)) +
  geom_point() +
  xlab("z = x^3") +
  ylab("y") +
  geom_smooth(method = "lm", se = FALSE)
curve_plt2
```
</center>
 
The process of
transforming predictors (and potentially combining multiple predictors in the process)
is known as *feature engineering*. In real data analysis
problems, you will need to rely on
a deep understanding of the problem---as well as the wrangling tools 
from previous chapters---to engineer useful new features that improve
predictive performance.

> Note that feature engineering
> is *part of tuning your model*, and as such you must not use your test data
> to evaluate the quality of the features you produce. You are free to use
> cross-validation, though!

## The other sides of regression

So far in this textbook we have used regression only in the context of
prediction. However, regression can also be seen as a method to understand and
quantify the effects of individual variables on a response / outcome of interest.
In the housing example from this chapter, beyond just using past data
to predict future sale prices, 
we might also be interested in describing the
individual relationships of house size and the number of bedrooms with house price,
quantifying how strong each of these relationships are, and assessing how accurately we
can estimate their magnitudes. And even beyond that, we may be interested in 
understanding whether the predictors *cause* changes in the price.
These sides of regression are well beyond the scope of this book; but
the material you have learned here should give you a foundation of knowledge
that will serve you well when moving to more advanced books on the topic.

## Additional resources
- The [`tidymodels` website](https://tidymodels.org/packages) is an excellent reference for more details on, and advanced usage of, the functions and packages in the past two chapters. Aside from that, it also has a [nice beginner's tutorial](https://www.tidymodels.org/start/) and [an extensive list of more advanced examples](https://www.tidymodels.org/learn/) that you can use to continue learning beyond the scope of this book. 
- [Modern Dive](https://moderndive.com/) is another textbook that uses the `tidyverse` / `tidymodels` framework. Chapter 6 complements the material in the current chapter well; it covers some slightly more advanced concepts than we do without getting mathematical. Give this chapter a read before moving on to the next reference. It is also worth noting that this book takes a more "explanatory" / "inferential" approach to regression in general (in Chapters 5, 6, and 10), which provides a nice complement to the predictive tack we take in the present book.
- [An Introduction to Statistical Learning](https://www.statlearning.com/) [-@james2013introduction] provides a great next stop in the process of learning about regression. Chapter 3 covers linear regression at a slightly more mathematical level than we do here, but it is not too large a leap and so should provide a good stepping stone. Chapter 6 discusses how to pick a subset of "informative" predictors when you have a data set with many predictors, and you expect only a few of them to be relevant. Chapter 7 covers regression models that are more flexible than linear regression models but still enjoy the computational efficiency of linear regression. In contrast, the KNN methods we covered earlier are indeed more flexible but become very slow when given lots of data.
