# Collaboration with version control {#version-control}

```{r 12-getting-started-with-version-control, echo = FALSE, message = FALSE, warning = FALSE}
library(magick)
library(magrittr)
library(knitr)

knitr::opts_chunk$set(message = FALSE,
                      echo = FALSE,
                      warning = FALSE,
                      fig.align = "center")
```

> *You mostly collaborate with yourself,
> and me-from-two-months-ago never responds to email.*
>
> --Mark T. Holder


## Overview

This chapter will introduce the concept of using version control systems
to track changes to a project over its lifespan, to share
and edit code in a collaborative team,
and to distribute the finished project to its intended audience.
This chapter will also introduce how to use
the two most common version control tools: Git \index{git} for local version control,
and GitHub \index{GitHub} for remote version control.
We will focus on the most common version control operations
used day-to-day in a standard data science project.
There are many user interfaces for Git; in this chapter
we will cover the Jupyter Git interface.

## Chapter learning objectives

By the end of the chapter, readers will be able to do the following:

- Describe what version control is and why data analysis projects can benefit from it.
- Create a remote version control repository on GitHub.
- Use Jupyter's Git version control tools for project versioning and collaboration:
  - Clone a remote version control repository to create a local repository.
  - Commit changes to a local version control repository.
  - Push local changes to a remote version control repository.
  - Pull changes from a remote version control repository to a local version control repository.
  - Resolve merge conflicts.
- Give collaborators access to a remote GitHub repository.
- Communicate with collaborators using GitHub issues.
- Use best practices when collaborating on a project with others.

## What is version control, and why should I use it?

Data analysis projects often require iteration
and revision to move from an initial idea to a finished product
ready for the intended audience.
Without deliberate and conscious effort towards tracking changes
made to the analysis, projects tend to become messy.
This mess can have serious, negative repercussions on an analysis project,
including results that your code cannot reproduce,
temporary files with snippets of ideas that are forgotten or
not easy to find, mind-boggling file names that make it unclear which is
the current working version of the file (e.g., `document_final_draft_final.txt`,
`to_hand_in_final_v2.txt`, etc.), and more.

Additionally, the iterative nature of data analysis projects
means that most of the time, the final version of the analysis that is
shared with the audience is only a fraction of what was explored during
the development of that analysis.
Changes in data visualizations and modeling approaches,
as well as some negative results, are often not observable from
reviewing only the final, polished analysis.
The lack of observability of these parts of the analysis development
can lead to others repeating things that did not work well,
instead of seeing what did not work well,
and using that as a springboard to new, more fruitful approaches.

Finally, data analyses are typically completed by a team of people
rather than a single person.
This means that files need to be shared across multiple computers,
and multiple people often end up editing the project simultaneously.
In such a situation, determining who has the latest version of the
project&mdash;and how to resolve conflicting edits&mdash;can be a real challenge.

*Version control* \index{version control} helps solve these challenges. Version control is the process
of keeping a record of changes to documents, including when the changes were
made and who made them, throughout the history of their development.  It also
provides the means both to view earlier versions of the project and to revert
changes.  Version control is most commonly used in software development, but
can be used for any electronic files for any type of project, including data
analyses.  Being able to record and view the history of a data analysis project
is important for understanding how and why decisions to use one method or
another were made, among other things.  Version control also facilitates
collaboration via tools to share edits with others and resolve conflicting
edits.  But even if you're working on a project alone, you should still use
version control.  It helps you keep track of what you've done, when you did it,
and what you're planning to do next!


To version control a project, you generally need two things:
a *version control system* \index{version control!system} and a *repository hosting service*. \index{version control!repository hosting}
The version control system is the software responsible
for tracking changes, sharing changes you make with others,
obtaining changes from others, and resolving conflicting edits.
The repository hosting service is responsible for storing a copy
of the version-controlled project online (a *repository*),
where you and your collaborators can access it remotely,
discuss issues and bugs, and distribute your final product.
For both of these items, there is a wide variety of choices.
In this textbook we'll use Git for version control,
and GitHub for repository hosting,
because both are currently the most widely used platforms.
In the
additional resources section at the end of the chapter,
we list many of the common version control systems
and repository hosting services in use today.

> **Note:** Technically you don't *have to* use a repository hosting service.
> You can, for example, version control a project
> that is stored only in a folder on your computer&mdash;never
> sharing it on a repository hosting service.
> But using a repository hosting service provides a few big benefits,
> including managing collaborator access permissions,
> tools to discuss and track bugs,
> and the ability to have external collaborators contribute work,
> not to mention the safety of having your work backed up in the cloud.
> Since most repository hosting services now offer free accounts,
> there are not many situations in which you wouldn't
> want to use one for your project.

## Version control repositories

Typically, when we put a data analysis project under version control,
we create two copies of the repository \index{repository} (Figure \@ref(fig:vc1-no-changes)).
One copy we use as our primary workspace where we create, edit, and delete files.
This copy is commonly referred to as \index{repository!local} the **local repository**. The local
repository most commonly exists on our computer or laptop, but can also exist within
a workspace on a server (e.g., JupyterHub).
The other copy is typically stored in a repository hosting service (e.g., GitHub), where
we can easily share it with our collaborators.
This copy is commonly referred to as \index{repository!remote} the **remote repository**.
\index{repository|see{version control}}

```{r vc1-no-changes, fig.align = "center", fig.cap = 'Schematic of local and remote version control repositories.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc1-no-changes.png") |>
  image_crop("3632x2000")
```

Both copies of the repository have a **working directory** \index{working directory}
where you can create, store, edit, and delete
files (e.g., `analysis.ipynb` in Figure \@ref(fig:vc1-no-changes)).
Both copies of the repository also maintain a full project history
(Figure \@ref(fig:vc1-no-changes)).  This history is a record of all versions of the
project files that have been created.  The repository history is not
automatically generated; Git must be explicitly told when to record
a version of the project.  These records are \index{git!commit} called **commits**. They
are a snapshot of the file contents as well
metadata about the repository at that time the record was created (who made the
commit, when it was made, etc.). In the local and remote repositories shown in
Figure \@ref(fig:vc1-no-changes), there are two commits represented as light blue
rectangles inside the "Repository History" sections.  Each commit can be identified by a
human-readable **message**, which you write when you make a commit, and a
**commit hash** that Git automatically adds for you.

The purpose of the message is to contain a brief, rich description
of what work was done since the last commit.
Messages act as a very useful narrative
of the changes to a project over its lifespan.
If you ever want to view or revert to an earlier version of the project,
the message can help you identify which commit to view or revert to.
In Figure \@ref(fig:vc1-no-changes), you can see two such messages,
one for each commit: `Created README.md` and `Added analysis draft`.

The hash \index{hash} is a string of characters consisting of about 40 letters and numbers.
The purpose of the hash is to serve as a unique identifier for the commit,
and is used by Git to index project history. Although hashes are quite long&mdash;imagine
having to type out 40 precise characters to view an old project version!&mdash;Git is able
to work with shorter versions of hashes. In Figure \@ref(fig:vc1-no-changes), you can see
two of these shortened hashes, one for each commit: `Daa29d6` and `884c7ce`.

## Version control workflows

When you work in a local version-controlled repository, there are generally three additional
steps you must take as part of your regular workflow. In addition to
just working on files&mdash;creating,
editing, and deleting files as you normally would&mdash;you must:

1. Tell Git when to make a commit of your own changes in the local repository.
2. Tell Git when to send your new commits to the remote GitHub repository.
3. Tell Git when to retrieve any new changes (that others made) from the remote GitHub repository.

In this section we will discuss all three of these steps in detail.

### Committing changes to a local repository {#commit-changes}

When working on files in your local version control
repository (e.g., using Jupyter) and saving your work, these changes will only initially exist in the
working directory of the local repository (Figure \@ref(fig:vc2-changes)).

```{r vc2-changes, fig.align = "center", fig.cap = 'Local repository with changes to files.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc2-changes.png") |>
  image_crop("3632x2000")
```

Once you reach a point that you want Git to keep a record
of the current version of your work, you need to **commit** \index{git!commit}
(i.e., snapshot) your changes. A prerequisite to this is telling Git which
files should be included in that snapshot. We call this step **adding** the
files to the **staging area**. \index{git!add}\index{git!staging area}\index{staging area|see{git}}
Note that the staging area is not a real physical location on your computer;
it is instead a conceptual placeholder for these files until they are committed.
The benefit of the Git version control system using a staging area is that you
can choose to commit changes in only certain files. For example,
in Figure \@ref(fig:vc-ba2-add), we add only the two files
that are important to the analysis project (`analysis.ipynb` and `README.md`)
and not our personal scratch notes for the project (`notes.txt`).

```{r vc-ba2-add, fig.align = "center", fig.cap = 'Adding modified files to the staging area in the local repository.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc-ba2-add.png") |>
  image_crop("3632x1200")
```

Once the files we wish to commit have been added
to the staging area, we can then commit those files to the repository history (Figure \@ref(fig:vc-ba3-commit)).
When we do this, we are required to include a helpful *commit message* to tell
collaborators (which often includes future you!) about the changes that were
made. In Figure \@ref(fig:vc-ba3-commit), the message is `Message about changes...`; in
your work you should make sure to replace this with an
informative message about what changed. It is also important to note here that
these changes are only being committed to the local repository's history.  The
remote repository on GitHub has not changed, and collaborators are not yet
able to see your new changes.

```{r vc-ba3-commit, fig.align = "center", fig.cap = "Committing the modified files in the staging area to the local repository history, with an informative message about what changed.", fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc-ba3-commit.png") |>
  image_crop("3632x1100")
```

### Pushing changes to a remote repository

Once you have made one or more commits that you want to share with your collaborators,
you need \index{git!push} to **push** (i.e., send) those commits back to GitHub (Figure \@ref(fig:vc5-push)). This updates
the history in the remote repository (i.e., GitHub) to match what you have in your
local repository. Now when collaborators interact with the remote repository, they will be able
to see the changes you made. And you can also take comfort in the fact that your work is now backed
up in the cloud!

```{r vc5-push, fig.align = "center", fig.cap = 'Pushing the commit to send the changes to the remote repository on GitHub.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc5-push.png") |>
  image_crop("3632x3000")
```

### Pulling changes from a remote repository

If you are working on a project with collaborators, they will also be making changes to files
(e.g., to the analysis code in a Jupyter notebook and the project's README file),
committing them to their own local repository, and pushing their commits to the remote GitHub repository
to share them with you. When they push their changes, those changes will only initially exist in
the remote GitHub repository and not in your local repository (Figure \@ref(fig:vc6-remote-changes)).

```{r vc6-remote-changes, fig.align = "center", fig.cap = 'Changes pushed by collaborators, or created directly on GitHub will not be automatically sent to your local repository.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc6-remote-changes.png") |>
  image_crop("3632x2000")
```

To obtain the new changes from the remote repository on GitHub, you will need
to **pull** \index{git!pull} those changes to your own local repository.  By pulling changes,
you synchronize your local repository to what is present on GitHub (Figure \@ref(fig:vc7-pull)).
Additionally, until you pull changes from the remote repository, you will not
be able to push any more changes yourself (though you will still be able to
work and make commits in your own local repository).

```{r vc7-pull, fig.align = "center", fig.cap = 'Pulling changes from the remote GitHub repository to synchronize your local repository.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/vc7-pull.png") |>
  image_crop("3632x2000")
```

## Working with remote repositories using GitHub

Now that you have been introduced to some of the key general concepts
and workflows of Git version control, we will walk through the practical steps.
There are several different ways to start using version control
with a new project. For simplicity and ease of setup,
we recommend creating a remote repository \index{repository!remote} first.
This section covers how to both create and edit a remote repository on \index{GitHub} GitHub.
Once you have a remote repository set up, we recommend **cloning** (or copying) that \index{git!clone}
repository to create a local repository in which you primarily work.
You can clone the repository either
on your own computer or in a workspace on a server (e.g., a JupyterHub server).
Section \@ref(local-repo-jupyter) below will cover this second step in detail.

### Creating a remote repository on GitHub

Before you can create remote repositories on GitHub,
you will need a GitHub account; you can sign up for a free account
at [https://github.com/](https://github.com/).
Once you have logged into your account, you can create a new repository to host
your project by clicking on the "+" icon in the upper right-hand
corner, and then on "New Repository," as shown in
Figure \@ref(fig:new-repository-01).

(ref:new-repository-01) New repositories on GitHub can be created by clicking on "New Repository" from the + menu.

```{r new-repository-01, fig.align = "center", fig.cap = '(ref:new-repository-01)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/new_repository_01.png") |>
  image_flop() |>
  image_crop("3632x1148") |>
  image_flop()
```

Repositories can be set up with a variety of configurations, including a name,
optional description,  and the inclusion (or not) of several template files.
One of the most important configuration items to choose is the visibility to the outside world,
either public or private. *Public* repositories \index{repository!public} can be viewed by anyone.
*Private* repositories \index{repository!private} can be viewed by only you. Both public and private repositories
are only editable by you, but you can change that by giving access to other collaborators.

To get started with a *public* repository having a template `README.md` file, take the
following steps shown in Figure \@ref(fig:new-repository-02):

1. Enter the name of your project repository. In the example below, we use `canadian_languages`. Most repositories follow a similar naming convention involving only lowercase letter words separated by either underscores or hyphens.
2. Choose an option for the privacy of your repository.
3. Select "Add a README file." This creates a template `README.md` file in your repository's root folder.
4. When you are happy with your repository name and configuration, click on the green "Create Repository" button.

```{r new-repository-02, fig.align = "center", fig.cap = 'Repository configuration for a project that is public and initialized with a README.md template file.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/new_repository_02.png") |>
  image_flop() |>
  image_crop("1700x2240+1000-100") |>
  image_flop()
```

A newly created public repository with a `README.md` template file should look something
like what is shown in Figure \@ref(fig:new-repository-03).

```{r new-repository-03, fig.align = "center", fig.cap = 'Respository configuration for a project that is public and initialized with a README.md template file.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/new_repository_03.png") |>
  image_flop() |>
  image_crop("3584x1700") |>
  image_flop()
```


### Editing files on GitHub with the pen tool

The pen tool \index{GitHub!pen tool} can be used to edit existing plain text files. When you click on
the pen tool, the file will be opened in a text box where you can use your
keyboard to make changes (Figures \@ref(fig:pen-tool-01) and \@ref(fig:pen-tool-02)).

```{r pen-tool-01, fig.align = "center", fig.cap = 'Clicking on the pen tool opens a text box for editing plain text files.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/pen-tool_01.png") |>
  image_flop() |>
  image_crop("3584x1500") |>
  image_flop()
```

```{r pen-tool-02, fig.align = "center", fig.cap = 'The text box where edits can be made after clicking on the pen tool.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/pen-tool_02.png") |>
  # image_flop() |>
  image_crop("3584x2000") # |>
# image_flop()
```

After you are done with your edits, they can be "saved" by *committing* your
changes. When you *commit a file* in a repository, the version control system
takes a snapshot of what the file looks like. As you continue working on the
project, over time you will possibly make many commits to a single file; this
generates a useful version history for that file. On GitHub, if you click the
green "Commit changes" button, \index{GitHub!commit} it will save the file and then make a commit
(Figure \@ref(fig:pen-tool-03)).

Recall from Section \@ref(commit-changes) that you normally have to add files
to the staging area before committing them. Why don't we have to do that when
we work directly on GitHub? Behind the scenes, when you click the green "Commit changes"
button, GitHub *is* adding that one file to the staging area prior to committing it.
But note that on GitHub you are limited to committing changes to only one file at a time.
When you work in your own local repository, you can commit
changes to multiple files simultaneously. This is especially useful when one
"improvement" to the project involves modifying multiple files.
You can also do things like run code when working in a local repository, which you cannot
do on GitHub. In general, editing on GitHub is reserved for small edits to plain text files.

```{r pen-tool-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Saving changes using the pen tool requires committing those changes, and an associated commit message.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/pen-tool_03.png") |>
  image_crop("3583x1500+1+500")
```

### Creating files on GitHub with the "Add file" menu

The "Add file" menu \index{GitHub!add file} can be used to create new plain text files and upload files
from your computer. To create a new plain text file, click the "Add file"
drop-down menu and select the "Create new file" option
(Figure \@ref(fig:create-new-file-01)).

```{r create-new-file-01, fig.align = "center", fig.cap = 'New plain text files can be created directly on GitHub.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/create-new-file_01.png") |>
  image_flop() |>
  image_crop("3584x1600") |>
  image_flop()
```

A page will open with a small text box for the file name to be entered, and a
larger text box where the desired file content text can be entered. Note the two
tabs, "Edit new file" and "Preview". Toggling between them lets you enter and
edit text and view what the text will look like when rendered, respectively
(Figure \@ref(fig:create-new-file-02)).
Note that GitHub understands and renders `.md` files \index{markdown} using a
markdown syntax very similar to Jupyter notebooks, so the "Preview" tab is especially helpful
for checking markdown code correctness.

```{r create-new-file-02, fig.align = "center", fig.cap = 'New plain text files require a file name in the text box circled in red, and file content entered in the larger text box (red arrow).', fig.retina = 2, out.width="100%"}
image_read("img/version-control/create-new-file_02.png") |>
  image_flop() |>
  image_crop("3584x1300") |>
  image_flop()
```

Save and commit your changes by clicking the green "Commit changes" button at the
bottom of the page (Figure \@ref(fig:create-new-file-03)).

```{r create-new-file-03, fig.align = "center", fig.cap = 'To be saved, newly created files are required to be committed along with an associated commit message.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/create-new-file_03.png") |>
  image_crop("3584x1500+1+500")
```

You can also upload files that you have created on your local machine by using
the "Add file" drop-down menu and selecting "Upload files"
(Figure \@ref(fig:upload-files-01)).
To select the files from your local computer to upload, you can either drag and
drop them into the gray box area shown below, or click the "choose your files"
link to access a file browser dialog. Once the files you want to upload have
been selected, click the green "Commit changes" button at the bottom of the
page (Figure \@ref(fig:upload-files-02)).

```{r upload-files-01, fig.align = "center", fig.cap = 'New files of any type can be uploaded to GitHub.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/upload-files_01.png") |>
  image_flop() |>
  image_crop("3584x1600") |>
  image_flop()
```

(ref:upload-files-02) Specify files to upload by dragging them into the GitHub website (red circle) or by clicking on "choose your files." Uploaded files are also required to be committed along with an associated commit message.

```{r upload-files-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:upload-files-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/upload-files_02.png") |>
  image_flop() |>
  image_crop("3584x2200") |>
  image_flop()
```

Note that Git and GitHub are designed to track changes in individual files.
**Do not** upload your whole project in an archive file (e.g., `.zip`). If you do,
then Git can only keep track of changes to the entire `.zip` file, which will not
be human-readable. Committing one big archive defeats the whole purpose of using
version control: you won't be able to see, interpret, or find changes in the history
of any of the actual content of your project!

## Working with local repositories using Jupyter {#local-repo-jupyter}

Although there are several ways to create and edit files on GitHub, they are
not quite powerful enough for efficiently creating and editing complex files,
or files that need to be executed to assess whether they work (e.g., files
containing code).  For example, you wouldn't be able to run an analysis written
with R code directly on GitHub.  Thus, it is useful to be able to connect the
remote repository that was created on GitHub to a local coding environment.  This
can be done by creating and working in a local copy of the repository.
In this chapter, we focus on interacting with Git via Jupyter using
the Jupyter Git extension. The Jupyter Git \index{git!Jupyter extension} extension
can be run by Jupyter on your local computer, or on a JupyterHub server.
We recommend reading Chapter \@ref(jupyter) to learn how
to use Jupyter before reading this chapter.

### Generating a GitHub personal access token

To send and retrieve work between your local repository
and the remote repository on GitHub,
you will frequently need to authenticate with GitHub
to prove you have the required permission.
There are several methods to do this,
but for beginners we recommend using the HTTPS method
because it is easier and requires less setup.
In order to use the HTTPS method,
GitHub requires you to provide a *personal access token*. \index{GitHub!personal access token}
A personal access token is like a password&mdash;so keep it a secret!&mdash;but it gives
you more fine-grained control over what parts of your account
the token can be used to access, and lets you set an expiry date for the authentication.
To generate a personal access token,
you must first visit [https://github.com/settings/tokens](https://github.com/settings/tokens),
which will take you to the "Personal access tokens" page in your account settings.
Once there, click "Generate new token" (Figure \@ref(fig:generate-pat-01)).
Note that you may be asked to re-authenticate with your username
and password to proceed.

(ref:generate-pat-01) The "Generate new token" button used to initiate the creation of a new personal access token. It is found in the "Personal access tokens" section of the "Developer settings" page in your account settings.

```{r generate-pat-01, fig.align = "center", fig.cap = '(ref:generate-pat-01)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/generate-pat_01.png")
```

You will be asked to add a note to describe the purpose for your personal access token.
Next, you need to select permissions for the token; this is where
you can control what parts of your account the token can be used to access.
Make sure to choose only those permissions that you absolutely require. In
Figure \@ref(fig:generate-pat-02), we tick only the "repo" box, which gives the
token access to our repositories (so that we can push and pull) but none of our other GitHub
account features. Finally, to generate the token, scroll to the bottom of that page
and click the green "Generate token" button (Figure \@ref(fig:generate-pat-02)).

(ref:generate-pat-02) Webpage for creating a new personal access token.

```{r generate-pat-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:generate-pat-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/generate-pat_02.png")
```

Finally, you will be taken to a page where you will be able to see
and copy the personal access token you just generated (Figure \@ref(fig:generate-pat-03)).
Since it provides access to certain parts of your account, you should
treat this token like a password; for example, you should consider
securely storing it (and your other passwords and tokens, too!) using a password manager.
Note that this page will only display the token to you once,
so make sure you store it in a safe place right away. If you accidentally forget to
store it, though, do not fret&mdash;you can delete that token by clicking the
"Delete" button next to your token, and generate a new one from scratch.
To learn more about GitHub authentication,
see the additional resources section at the end of this chapter.

(ref:generate-pat-03) Display of the newly generated personal access token.

```{r generate-pat-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:generate-pat-03)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/generate-pat_03.png")
```

### Cloning a repository using Jupyter

*Cloning* a \index{git!clone} remote repository from GitHub
to create a local repository results in a
copy that knows where it was obtained from so that it knows where to send/receive
new committed edits. In order to do this, first copy the URL from the HTTPS tab
of the Code drop-down menu on GitHub (Figure \@ref(fig:clone-02)).

(ref:clone-02) The green "Code" drop-down menu contains the remote address (URL) corresponding to the location of the remote GitHub repository.

```{r clone-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:clone-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/clone_02.png") |>
  image_crop("3584x1050")
```

Open Jupyter, and click the Git+ icon on the file browser tab
(Figure \@ref(fig:clone-01)).

```{r clone-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The Jupyter Git Clone icon (red circle).', fig.retina = 2, out.width="100%"}
image_read("img/version-control/clone_01.png") |>
  image_crop("2400x1300+1")
```

Paste the URL of the GitHub project repository you
created and click the blue "CLONE" button (Figure \@ref(fig:clone-03)).

```{r clone-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Prompt where the remote address (URL) corresponding to the location of the GitHub repository needs to be input in Jupyter.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/clone_03.png") |>
  image_crop("2400x1430+1")
```

On the file browser tab, you will now see a folder for the repository.
Inside this folder  will be all the files that existed on GitHub (Figure \@ref(fig:clone-04)).

```{r clone-04, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Cloned GitHub repositories can been seen and accessed via the Jupyter file browser.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/clone_04.png") |>
  image_crop("2400x1200+1")
```

### Specifying files to commit
Now that you have cloned the remote repository from GitHub to create a local repository,
you can get to work editing, creating, and deleting files.
For example, suppose you created and saved a new file (named `eda.ipynb`) that you would
like to send back to the project repository on GitHub (Figure \@ref(fig:git-add-01)).
To "add" this modified file to the staging area (i.e., flag that this is a
file whose changes we would like to commit), click the Jupyter Git extension
icon on the far left-hand side of Jupyter (Figure \@ref(fig:git-add-01)).

```{r git-add-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Jupyter Git extension icon (circled in red).', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_add_01.png") |>
  image_crop("3584x1200")
```

This opens the Jupyter Git graphical user interface pane. Next,
click the plus sign (+) beside the file(s) that you want to "add"  \index{git!add}
(Figure \@ref(fig:git-add-02)). Note that because this is the
first change for this file, it falls under the "Untracked" heading.
However, next time you edit this file  and want to add the changes,
you will find it under the "Changed" heading.

You will also see an `eda-checkpoint.ipynb` file under the "Untracked" heading.
This is a temporary "checkpoint file" created by Jupyter when you work on `eda.ipynb`.
You generally do not want to add auto-generated files to Git repositories;
only add the files you directly create and edit.

(ref:git-add-02) `eda.ipynb` is added to the staging area via the plus sign (+).

```{r git-add-02, fig.align = "center", fig.cap = '(ref:git-add-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_add_02.png") |>
  image_crop("3584x1200")
```

Clicking the plus sign (+) moves the file from the "Untracked" heading to the "Staged" heading,
so that Git knows you want a snapshot of its current state
as a commit (Figure \@ref(fig:git-add-03)).
Now you are ready to "commit" the changes.
Make sure to include a (clear and helpful!) message about what was changed
so that your collaborators (and future you) know what happened in this commit.

(ref:git-add-03) Adding `eda.ipynb` makes it visible in the staging area.

```{r git-add-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:git-add-03)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_add_03.png") |>
  image_crop("3584x1200")
```

### Making the commit

To snapshot the changes with an associated commit message,
you must put a message in the text box at the bottom of the Git pane
and click on the blue "Commit" button (Figure \@ref(fig:git-commit-01)). \index{git!commit}
It is highly recommended to write useful and meaningful messages about what
was changed. These commit messages, and the datetime stamp for a given
commit, are the primary means to navigate through the project's history in the
 event that you need to view or retrieve a past version of a file, or
revert your project to an earlier state.
When you click the "Commit" button for the first time, you will be prompted to
enter your name and email. This only needs to be done once for each machine
you use Git on.

```{r git-commit-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'A commit message must be added into the Jupyter Git extension commit text box before the blue Commit button can be used to record the commit.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_commit_01.png")
```

After "committing" the file(s), you will see there are 0 "Staged" files.
You are now ready to push your changes
to the remote repository on GitHub (Figure \@ref(fig:git-commit-03)).

```{r git-commit-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'After recording a commit, the staging area should be empty.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_commit_03.png") |>
  image_crop("3584x1500")
```

### Pushing the commits to GitHub

To send the committed changes back to the remote repository on
GitHub, you need to *push* them. \index{git!push} To do this,
click on the cloud icon with the up arrow on the Jupyter Git tab
(Figure \@ref(fig:git-push-01)).

(ref:git-push-01) The Jupyter Git extension "push" button (circled in red).

```{r git-push-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:git-push-01)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_push_01.png") |>
  image_crop("3584x1500")
```

You will then be prompted to enter your GitHub username
and the personal access token that you generated
earlier (not your account password!). Click
the blue "OK" button to initiate the push (Figure \@ref(fig:git-push-02)).

```{r git-push-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Enter your Git credentials to authorize the push to the remote repository.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_push_02.png") |>
  image_crop("3584x1900")
```

If the files were successfully pushed to the project repository on
GitHub, you will be shown a success message (Figure \@ref(fig:git-push-03)).
Click "Dismiss" to continue working in Jupyter.

```{r git-push-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The prompt that the push was successful.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_push_03.png") |>
  image_crop("3584x1900")
```

If you visit the remote repository on GitHub,
you will see that the changes now exist there too
(Figure \@ref(fig:git-push-04))!

```{r git-push-04, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The GitHub web interface shows a preview of the commit message, and the time of the most recently pushed commit for each file.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_push_04.png") |>
  image_crop("3584x1900")
```

## Collaboration

### Giving collaborators access to your project

As mentioned earlier, GitHub allows you to control who has access to your
project. The default of both public and private projects are that only the
person who created the GitHub \index{GitHub!collaborator access} repository has permissions to create, edit and
delete files (*write access*). To give your collaborators write access to the
projects, navigate to the "Settings" tab (Figure \@ref(fig:add-collab-01)).

(ref:add-collab-01) The "Settings" tab on the GitHub web interface.

```{r add-collab-01, fig.pos = "H", out.extra="",  fig.align = "center", fig.cap = '(ref:add-collab-01)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/add_collab_01.png") |>
  image_crop("3584x1250")
```

Then click "Manage access" (Figure \@ref(fig:add-collab-02)).

(ref:add-collab-02) The "Manage access" tab on the GitHub web interface.

```{r add-collab-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:add-collab-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/add_collab_02.png") |>
  image_crop("3584x1200")
```

Then click the green "Invite a collaborator" button (Figure \@ref(fig:add-collab-03)).

(ref:add-collab-03) The "Invite a collaborator" button on the GitHub web interface.

```{r add-collab-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:add-collab-03)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/add_collab_03.png") |>
  image_crop("3584x2200")
```

Type in the collaborator's GitHub username or email,
and select their name when it appears (Figure \@ref(fig:add-collab-04)).

```{r add-collab-04, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = "The text box where a collaborator's GitHub username or email can be entered.", fig.retina = 2, out.width="100%"}
image_read("img/version-control/add_collab_04.png") |>
  image_crop("3584x1250")
```

Finally, click the green "Add collaborator to this repository" button (Figure \@ref(fig:add-collab-05)).

```{r add-collab-05, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The confirmation button for adding a collaborator to a repository on the GitHub web interface.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/add_collab_05.png") |>
  image_crop("3584x1250")
```

After this, you should see your newly added collaborator listed under the
"Manage access" tab. They should receive an email invitation to join the
GitHub repository as a collaborator. They need to accept this invitation
to enable write access.

### Pulling changes from GitHub using Jupyter

We will now walk through how to use the Jupyter Git extension tool to pull changes
to our `eda.ipynb` analysis file that were made by a collaborator
(Figure \@ref(fig:git-pull-00)).

```{r git-pull-00, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The GitHub interface indicates the name of the last person to push a commit to the remote repository, a preview of the associated commit message, the unique commit identifier, and how long ago the commit was snapshotted.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_pull_00.png") |>
  image_crop("3584x1600")
```

You can tell Git to "pull" by \index{git!pull} clicking on the cloud icon with
the down arrow in Jupyter (Figure \@ref(fig:git-pull-01)).

```{r git-pull-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The Jupyter Git extension clone button.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_pull_01.png") |>
  image_crop("3584x1430")
```

Once the files are successfully pulled from GitHub, you need to click "Dismiss"
to keep working (Figure \@ref(fig:git-pull-02)).

```{r git-pull-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'The prompt after changes have been successfully pulled from a remote repository.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_pull_02.png") |>
  image_crop("3584x1450")
```

And then when you open (or refresh) the files whose changes you just pulled,
you should be able to see them (Figure \@ref(fig:git-pull-03)).

(ref:git-pull-03) Changes made by the collaborator to `eda.ipynb` (code highlighted by red arrows).

```{r git-pull-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:git-pull-03)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_pull_03.png") |>
  image_crop("3584x1450")
```

It can be very useful to review the history of the changes to your project. You
can do this directly in Jupyter by clicking "History" in the Git tab
(Figure \@ref(fig:git-pull-04)).

```{r git-pull-04, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Version control repository history viewed using the Jupyter Git extension.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/git_pull_04.png") |>
  image_crop("3584x1600")
```

It is good practice to pull any changes at the start of *every* work session
before you start working on your local copy.
If you do not do this,
and your collaborators have pushed some changes to the project to GitHub,
then you will be unable to push your changes to GitHub until you pull.
This situation can be recognized by the error message
shown in Figure \@ref(fig:merge-conflict-01).

```{r merge-conflict-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Error message that indicates that there are changes on the remote repository that you do not have locally.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/merge_conflict_01.png") |>
  image_crop("3584x1450")
```

Usually, getting out of this situation is not too troublesome. First you need
to pull the changes that exist on GitHub that you do not yet have in the local
repository.  Usually when this happens, Git can automatically merge the changes
for you, even if you and your collaborators were working on different parts of
the same file!

If, however, you and your collaborators made changes to the same line of the
same file, Git will not be able to automatically merge the changes&mdash;it will
not know whether to keep your version of the line(s), your collaborators
version of the line(s), or some blend of the two. When this happens, Git will
tell you that you have a merge conflict in certain file(s) (Figure \@ref(fig:merge-conflict-03)).

```{r merge-conflict-03, fig.align = "center", fig.cap = 'Error message that indicates you and your collaborators made changes to the same line of the same file and that Git will not be able to automatically merge the changes.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/merge_conflict_03.png") |>
  image_crop("3584x1450")
```

### Handling merge conflicts

To fix the merge conflict, \index{git!merge conflict} you need to open the offending file
in a plain text editor and look for special marks that Git puts in the file to
tell you where the merge conflict occurred (Figure \@ref(fig:merge-conflict-04)).

```{r merge-conflict-04, fig.align = "center", fig.cap = 'How to open a Jupyter notebook as a plain text file view in Jupyter.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/merge_conflict_04.png") |>
  image_crop("3584x1200")
```

The beginning of the merge
conflict is preceded by `<<<<<<< HEAD` and the end of the merge conflict is
marked by `>>>>>>>`. Between these markings, Git also inserts a separator
(`=======`). The version of the change before the separator is your change, and
the version that follows the separator was the change that existed on GitHub.
In Figure \@ref(fig:merge-conflict-05), you can see that in your local repository
there is a line of code that calls `scale_color_manual` with three color values (`deeppink2`, `cyan4`, and `purple1`).
It looks like your collaborator made an edit to that line too, except with different colors (to `blue3`, `red3`, and `black`)!

```{r merge-conflict-05, fig.align = "center", fig.cap = 'Merge conflict identifiers (highlighted in red).', fig.retina = 2, out.width="100%"}
image_read("img/version-control/merge_conflict_05.png") |>
  image_crop("3584x1400")
```

Once you have decided which version of the change (or what combination!) to
keep, you need to use the plain text editor to remove the special marks that
Git added (Figure \@ref(fig:merge-conflict-06)).

```{r merge-conflict-06, fig.align = "center", fig.cap = 'File where a merge conflict has been resolved.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/merge_conflict_06.png") |>
  image_crop("3584x1400")
```

The file must be saved, added to the staging area, and then committed before you will be able to
push your changes to GitHub.

### Communicating using GitHub issues

When working on a project in a team, you don't just want a historical record of who changed
what file and when in the project&mdash;you also want a record of decisions that were made,
ideas that were floated, problems that were identified and addressed, and all other
communication surrounding the project. Email and messaging apps are both very popular for general communication, but are not
designed for project-specific communication: they both generally do not have facilities for organizing conversations by project subtopics,
searching for conversations related to particular bugs or software versions, etc.

GitHub *issues* \index{GitHub!issues} are an alternative written communication medium to email and
messaging apps, and were designed specifically to facilitate project-specific
communication. Issues are *opened* from the "Issues" tab on the project's
GitHub page, and they persist there even after the conversation is over and the issue is *closed* (in
contrast to email, issues are not usually deleted). One issue thread is usually created
per topic, and they are easily searchable using GitHub's search tools. All
issues are accessible to all project collaborators, so no one is left out of
the conversation. Finally, issues can be set up so that team members get email
notifications when a new issue is created or a new post is made in an issue
thread. Replying to issues from email is also possible. Given all of these advantages,
 we highly recommend the use of issues for project-related communication.

To open a GitHub issue,
first click on the "Issues" tab (Figure \@ref(fig:issue-01)).

(ref:issue-01) The "Issues" tab on the GitHub web interface.

```{r issue-01, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:issue-01)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/issue_01.png") |>
  image_crop("3584x1700")
```

\newpage

Next click the "New issue" button (Figure \@ref(fig:issue-02)).

(ref:issue-02) The "New issue" button on the GitHub web interface.

```{r issue-02, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:issue-02)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/issue_02.png") |>
  image_crop("3584x1250")
```

Add an issue title (which acts like an email subject line), and then put the
body of the message in the larger text box. Finally, click "Submit new issue"
to post the issue to share with others (Figure \@ref(fig:issue-03)).

```{r issue-03, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Dialog boxes and submission button for creating new GitHub issues.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/issue_03.png") |>
  image_crop("3584x2200")
```

You can reply to an issue that someone opened by adding your written response to
the large text box and clicking comment (Figure \@ref(fig:issue-04)).

```{r issue-04, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = 'Dialog box for replying to GitHub issues.', fig.retina = 2, out.width="100%"}
image_read("img/version-control/issue_04.png") |>
  image_crop("3584x2000")
```

When a conversation is resolved, you can click "Close issue".
The closed issue can be later viewed by clicking the "Closed" header link
in the "Issue" tab (Figure \@ref(fig:issue-06)).

(ref:issue-06) The "Closed" issues tab on the GitHub web interface.

```{r issue-06, fig.pos = "H", out.extra="", fig.align = "center", fig.cap = '(ref:issue-06)', fig.retina = 2, out.width="100%"}
image_read("img/version-control/issue_06.png") |>
  image_crop("3584x1900")
```

## Exercises

Practice exercises for the material covered in this chapter
can be found in the accompanying
[worksheets repository](https://worksheets.datasciencebook.ca)
in the "Collaboration with version control" row.
You can launch an interactive version of the worksheet in your browser by clicking the "launch binder" button.
You can also preview a non-interactive version of the worksheet by clicking "view worksheet."
If you instead decide to download the worksheet and run it on your own machine,
make sure to follow the instructions for computer setup
found in Chapter \@ref(setup). This will ensure that the automated feedback
and guidance that the worksheets provide will function as intended.

## Additional resources {#vc-add-res}

Now that you've picked up the basics of version control with Git and GitHub,
you can expand your knowledge through the resources listed below:

- GitHub's [guides website](https://guides.github.com/) and [YouTube
  channel](https://www.youtube.com/githubguides), and [*Happy Git and GitHub
  for the useR*](https://happygitwithr.com/) are great resources to take the next steps in
  learning about Git and GitHub.
- [Good enough practices in scientific
  computing](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510#sec014)
  [@wilson2014best] provides more advice on useful workflows and "good enough"
  practices in data analysis projects.
- In addition to [GitHub](https://github.com), there are other popular Git
  repository hosting services such as [GitLab](https://gitlab.com) and
  [BitBucket](https://bitbucket.org). Comparing all of these options is beyond
  the scope of this book, and until you become a more advanced user, you are
  perfectly fine to just stick with GitHub. Just be aware that you have options!
- GitHub's [documentation on creating a personal access
  token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token)
  and the *Happy Git and GitHub for the useR*
  [personal access tokens chapter](https://happygitwithr.com/https-pat.html) are both
  excellent additional resources to consult if you need additional help
  generating and using personal access tokens.
